### 2020-05-05 notes.Rmd

---
title: "notes"
author: "Gayoung Park"
date: "5/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

I moved the leftover codes from paper, just in case I need them later.

```{r}

cv_plot_site %>% 
pivot_longer(cols = -sitename, names_to = "group") %>% 
ggplot(aes(x = sitename, y = value, fill = as.factor(group))) + 
geom_bar(stat="identity", position=position_dodge())

## I coded new one with N=artifact. 
```

Computing Pearson's correlation 
```{r}

#simpler way to calculate Correlataion, you can use
SCOR <- cor(DF)

#a scatterplot matrix of Correlation
ALL<-pairs(DF)

# correlation per phase
cor2 <- 
df_sitename %>% 
  filter(SPstage.Stage == 2) %>% 
  select(-sitename, -SPstage.Stage) %>% 
  cor(.)

ACOR2 <- pairs(cor2)

cor3 <- 
df_sitename %>% 
  filter(SPstage.Stage == 3) %>% 
  select(-sitename, -SPstage.Stage) %>% 
  cor(.)

ACOR3 <- pairs(cor3)

# Plot for SCOR
library(corrplot)

par(mfrow=c(1,3))
# correlation for all
cplotall <- corrplot(SCOR, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45, main = "Correlatoin for ALL")
# for phase 2
cplot2 <- corrplot(cor2, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45, main = "Correlatoin for Phase 2")
# for phase 3
cplot3 <- corrplot(cor3, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45, main = "Correlatoin for Phase 3")

```

```{R}

#simpler way to calculate Correlataion, you can use
SYG6_cor <- cor(SYG6_DF)

#a scatterplot matrix of Correlation
pairs(SYG6_DF)

cplot_SYG6 <- corrplot(SYG6_cor, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)

#simpler way to calculate Correlataion, you can use
SYG1_2cor <- cor(SYG1_2DF)

#a scatterplot matrix of Correlation
pairs(SYG1_2DF)

cplot_SYG1_2 <- corrplot(SYG1_2cor, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)

#simpler way to calculate Correlataion, you can use
SYG1_1cor <- cor(SYG1_1DF)

#a scatterplot matrix of Correlation
pairs(SYG1_1DF)

cplot_SYG1_1 <- corrplot(SYG1_1cor, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)

par(mfrow=c(1,3))
cplot_SYG6_ <- corrplot(SYG6_cor, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45, title = "SYG6")
cplot_SYG1_1 <- corrplot(SYG1_1cor, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45, title = "SYG1_1")
cplot_SYG1_2 <- corrplot(SYG1_2cor, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45, title = "SYG1_2")

```

```{r}
#I'm working on how to read oulines and how to use them for anlaysis. We may not use this for the research.
# reading landmarks and outline in tps, code from Morpho package, adjust readallTPS function.

readalltps = function(file) 
{
   out = list()
   exOut <- FALSE
  noutline <- 0
  input <- readLines(file)
   LM <- grep("LM=", input)
  LMstring <- input[LM]
   nLM <- sapply(LMstring, strsplit, split = "=")
  nLM <- unlist(nLM)
   nLM <- as.integer(nLM[-which(nLM == "LM")])
   nobs <- length(nLM)
   ID <- grep("ID=", input)
   IDstring <- input[ID]
    nID <- sapply(IDstring, function(x) {
       x <- gsub("=", "_", x)
   })
   nID <- gsub(" ", "", unlist(nID))
   out$ID <- nID
   outline <- grep("CURVES=", input)
   if (length(outline) > 0) {
       exOut <- TRUE
       outlinestring <- input[outline]
      noutline <- sapply(outlinestring, strsplit, split = "=")
      noutline <- unlist(noutline)
       noutline <- as.integer(noutline[-which(noutline == "CURVES")])
       LMoutline <- (sapply(outline, function(x) {
           x <- max(which(LM < x))  })) }
  LMdata <- list()
   for (i in 1:nobs) {
       if (nLM[i] > 0) {
           LMdata[[i]] <- as.numeric(unlist(strsplit(unlist(input[c((LM[i] + 
               1):(LM[i] + nLM[i]))]), split = " ")))
           LMdata[[i]] <- matrix(LMdata[[i]], nLM[i], 2, byrow = TRUE)
        }
      else LMdata[[i]] <- NA }
    names(LMdata) <- nID
    out$LM <- LMdata
    if (exOut) {
        outlineData <- list()
       for (i in 1:nobs) {
            if (i %in% LMoutline) {
                i1 <- grep(i, LMoutline)
               outlinetmp <- list()
               ptr <- outline[i1] + 1
               j <- 1
                while (j <= noutline[i1]) {
                 tmpnr <- as.integer(unlist(strsplit(input[ptr], 
                   split = "="))[2])
                  outlinetmp[[j]] <- matrix(as.numeric(unlist(strsplit(input[(ptr + 
                  1):(ptr + tmpnr)], split = " "))), tmpnr, 
                   2, byrow = TRUE)
                 ptr <- ptr + tmpnr + 1
                 j <- j + 1  }
              outlineData[[i]] <- outlinetmp  }
          else outlineData[[i]] <- NA }
       names(outlineData) <- nID
       out$outlines <- outlineData }
   cat(paste("Read", nobs, "datasets with", sum(noutline), "outlines\n"))
    return(out)}

#try to read all data(landmarks and outlines) from one file
 lm_trial <- readalltps(here("analysis/data/raw_data/A_BG_2.tps"))

#read all files from raw_data
 tps_files <- list.files(here("analysis/data/raw_data"), full.names = TRUE)
lm_landmarks <- map(tps_files, readalltps)
 names(lm_landmarks) <- list.files(here("analysis/data/raw_data"))

lm_landmarks = c()
for (i in 1:length(files)) {
  filename <- paste(datadir, files[i], sep = "")
  print(filename)
  landmarks <- readalltps(filename)
 lm_landmarks <- c(lm_landmarks, landmarks)}

```