### 2020-05-05 analysis/paper/paper.Rmd
refine some plots of CVs, sites, phases

---
title: "Title Goes Here"
author:
  -gayoung Park
  - Author Two
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::html_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

```{r, setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/"
)

#library(CTtps) # Or use devtools::load_all('.', quiet = T) if your code is in script files, rather than as functions in the `/R` diretory

library(here)
library(tidyverse)
```

# Introduction

1. What are our research questions?
- Understanding technological transition during the Late Pleistocene in Korea base on Cultural transmission. 
- What are the social contexts led to the technological transition

2. What is our archaeological background?
-SP was started in Korean Peninsula -> spread to Japan
-Why SP important, Why we care
-Current explantion of SP dispersal: migration vs in situ, in between
 The evidence of the two main models: proportion of SP in assemblage, Y chromosome (two groups of people), composition of assemblage (blade to flake ratio)
 -> no explanation about any mechanism/process/driver for the new technology
 -> lack of details of the lithic technology, human behavior

3. What is our conceptual framework?
- Evolutionary concepts that links stone artifacts to human behavior
- Cultural Evolution : the idea that cultural change constitutes an evolutionary process
- Same general concept of biological evolution: genetic change in population that is inherited over generations. Apply this concept to artifacts or cultural phenomena
- In this paper, I will use Cultural Transmission 
- Cultural Transmssion: way that people learn and pass on information
- Transmission biases
- similar study (CT in stone artifacts) : Mesoudi&O'Brien, Bettinger&Eerkens, Garvey

4. What are our methods?
- Framework: Bettinger&Eerkens
- Two models: Guided variation and Model-based bias =>  Multi-point Origin (in situ or in between one) and Single-point Origin (migration model)
- Test by Correlation and Variation (Pearson's r, PCA, Coefficient of Variation) 
- data collection, sites, artifact #
- attributes: landmarks (tps software)

5. What are our results? 
- overall change through time
- correlation: postively correlated(?). We are using PCA too
- PCA: 2,3 phases are highly overlapped, high redundancy. The individual phase are scattered, low redundancy. Variables of Dimension 1 is positively correlated
- CV:values are around 20(ML)-34(SL). Compared to 2nd phase, 3rd phase has slightly lower variation.

6. What is the answer to our questions??
- low correlation + high variation? High correlation + low variation?
- What are the criteria for correlation and variation
- How was SP spread in Korea, did they have single orgin or multiple origins?
- What are the impact of combinaton with blade technology

7. What is limitation of this study?
- Maybe applying CV for SP is not ideal. It can be depending on individual knapper's skill and raw materials
- CV of one site (SYG6) - no big difference with the value of all data
- CV: lithis vs pottery

- couldn't include thickness and weight

** Expectation
- The main measurements are correlation(+PCA) and CV.
- The set of the two measurements would be like: 
1) lower CV and higher correlation between sites, in this case, we will think that SPs were the result of model based bias and possibly had single origin. 
2) higher CV and lower correlation between sites, for this case, we will thank that SPs were the result of guided variation and possibly had multiple origins. 
- We have three phases, but there are only two artifacts for the first phase. There are more artifacts and sites for the first phase but not eligible for this research because we gathered complete or almost complete shape of SPs. We excluded broken pieces. Therefore, we focus on the change from the 2nd phase to 3rd phase for understanding the process of SP's dispersal in the Korean Peninsula.
- Change from the 2nd to 3rd
1) less correlated: the two groups were not very related/locally developed 
2) highly correlated: the two groups were related, had single source and then spread to other places
- Site by site
- Regional relationship? Province/cities

# Background

# Methods

Reading tps data and phase information of each artifact 
```{r results='hide', include=FALSE}
#reading landmarks in tps, code from https://gist.github.com/mrdwab/2062329

read.tps = function(data) {
  # Reads the .tps file format produced by TPSDIG 
  # (http://life.bio.sunysb.edu/morph/ into a single data frame
  # USAGE: R> read.tps("filename.tps")
  a = readLines(data) # so we can do some searching and indexing
  LM = grep("LM", a) # find the line numbers for LM
  ID.ind = grep("ID", a) # find the line numbers for ID
  # and the ID values, SCALE values, and image names
  ID = gsub("(ID=)(.*)", "\\2", grep("ID", a, value=T)) 
  SCALE = gsub("(SCALE=)(.*)", "\\2", grep("SCALE", a, value=T)) 
  images = basename(gsub("(IMAGE=)(.*)", "\\2", a[ID.ind - 1]))
  # FOR EACH LOOP   
  skip = LM # set how many lines to skip
  # and how many rows to read
  nrows = as.numeric(gsub("(LM=)(.*)", "\\2", grep("LM", a, value=T)))
  l = length(LM) # number of loops we want

  landmarks = vector("list", l) # create an empty list

  for (i in 1:l) {
    landmarks[i] = list(data.frame(
      read.table(file = data, header = F, skip = LM[i],
                 nrows = nrows[i], col.names = c("X", "Y")),
      IMAGE = as.character(images[i]),
      ID = ID[i],
      SCALE = SCALE[i]))
  }
  do.call(rbind, landmarks) # rbind the list items into a data.frame
}

#need to figure out how to read the file inside the project
landmarks <- read.tps(here("analysis/data/raw_data/BG_0_2.tps"))

#getting info for phase of each site
SPstage <- read_csv(here("analysis/data/SPstage.csv"))

#Compare multiple artifacts 
#Open multiple tps files. This will need to be changed to read files inside the project
datadir <- here("/analysis/data/raw_data/")
files <- dir(datadir, pattern = "*.tps$|*.TPS$")
```

Computing size attributes of the points by calculating distance between landmakrs
```{r}

#Understanding each landmark
#ML: maximum length, tip to bottom, perpendicular to the length axis 1- 7
#BL: body length: tip to the closest wing, perpendicular to the length axis 1- 3 OR 1-10 depending on the artifact
#TL: Tang length, the closest wing from the tip to bottom, perpendicular to the length axis 9 - 7 depending on the artifact
#SL: max stem length, perpendicular to the length axis. Closer tang curve’s middle point from the tip to the most distant point of the basal end  9-7 depending on the artifact
#MW: mid width: dimension from margin to margin at the mid-point of the length 2 - 11 axis, perpendicular to the width axis.
#TW: tang width: dimension between each wing, perpendicular to the width  axis  3 - 10
#SW: stem width: width of the basal end of the point, 5mm above the end  5 - 8

#distance between landmarks(euclidean)

landmark_dist = function(landmarks, pt1, pt2) {
  sqrt(
    (landmarks$X[pt1] - landmarks$X[pt2])^2 + 
      (landmarks$Y[pt1] - landmarks$Y[pt2])^2
  ) * as.numeric(levels(landmarks$SCALE))
}

# distance between two y axis 
vertical_dist = function(landmarks, pt1, pt2) {
  d = abs(landmarks$Y[pt1] - landmarks$Y[pt2])
  d * as.numeric(levels(landmarks$SCALE))
}

#Maximum length of sp: ML, distance between landmark 1-7
ML <- landmark_dist(landmarks, 1, 7)

#Body length of sp is 1-3 or 1-10, depending on each artifact.
bl_dist = function(landmarks) {
  d1 = landmark_dist(landmarks, 1, 3)
  d2 = landmark_dist(landmarks, 1, 10)
  max(d1, d2)
}

BL <- bl_dist(landmarks)

# Tang length of sp: TL, distance between landmark 10-7 or 3-7, (y axis)
tl_dist <- function(landmarks) {
  t1 = vertical_dist(landmarks, 10, 7)
  t2 = vertical_dist(landmarks, 3, 7)
  max(t1, t2)
}

TL <- tl_dist

# Maximum length of stem: SL, distance between landmark 9-7 or 4-7 (y axis)
sl_dist <- function(landmarks) {
  s1 = vertical_dist(landmarks, 9, 7)
  s2 = vertical_dist(landmarks, 4, 7)
  max(s1, s2)
}

SL <- sl_dist

# Mid width: MW, distance between landmark 2-11
MW <- landmark_dist(landmarks, 2, 11)

# Maximum width of Tang: TW, distance between landmark 3-10
TW <- landmark_dist(landmarks, 3, 10)

# Stem width: SW, distance between landmark 5-8
SW <- landmark_dist(landmarks, 5, 8)

```

Gather attributes from  all artifacts
```{r}

ML = c()
BL = c()
TL = c()
SL = c()
MW = c()
TW = c()
SW = c()
for (i in 1:length(files)) {
  filename <- paste(datadir, files[i], sep = "")
 # print(filename)
  landmarks <- read.tps(filename)
  ML <- c(ML, landmark_dist(landmarks, 1, 7))
  BL <- c(BL, bl_dist(landmarks))
  TL <- c(TL, tl_dist(landmarks))
  SL <- c(SL, sl_dist(landmarks))
  MW <- c(MW, landmark_dist(landmarks, 2, 11))
  TW <- c(TW, landmark_dist(landmarks, 3, 10))
  SW <- c(SW, landmark_dist(landmarks, 5, 8))
}

#list of all attributes
df <- data.frame(ML, BL, TL, SL, MW, TW, SW, files, SPstage$Stage)

#trim the name of each artifact.
df_sitename <- df %>% 
  mutate (files = str_remove_all(files, "_.tps")) %>% 
  separate (files, into = c("sitename", "files"), sep = "_") %>% 
  select (-files)

#list of all attributes without ID
DF <- df[,1:7]
```

Box plot for each attributes
```{r}
box1 <- boxplot(DF, main = "Variation for each attributes", col = "gold")
```

This boxplot indicates the overview of each attribute. We can see that ML and BL have wider variation compared to other attribtues. Since TL has relativelly less variable, ML variation is more depending on BL. Throughout time and region, dimension of SW is consistant.

Computing Coefficient or variation
```{r}
# Coefficient of variation (CV)

cv <- function(x, ... ) sd(x, ...) / mean(x, ...) *100

# we've got DF a data frame with cols as variables
# rows and specimens, and values as dimensions
# need to add a column of site ID
cvmap <- map(DF, cv)

#Same as map but better view for summary
map_df(DF, cv)
box2 <- boxplot(cvmap) ##Is there any better way to show this data??

```
The coefficient of variation (CV) is the ratio of the standard deviation devided by mean. It shows the extent of variability from the mean. 
CV for all attributes are distributed from 23-31%. Interestingly, SL and SW have high CV values (36.3 and 32.0, respectively). These attributes show lower variation on boxplot. <= need to think about

Make a table of CVs for all variable grouped by site

```{r}
cv_by_site_df  <- 
df_sitename %>% 
  select(-SPstage.Stage) %>% 
  group_by(sitename) %>% 
  add_tally() %>% 
  filter( n > 1) %>% 
  select(-n) %>% 
 nest(-sitename) %>% 
  mutate(cv_by_site = map(data, ~map_df(.x, cv)))

cv_by_site_df_unnest <-   
cv_by_site_df %>% 
  unnest(cv_by_site)

```

Make a plot for CV by site
```{r}
cv_plot_site <- cv_by_site_df_unnest %>% 
  select(-data)

library(tidyr)
cv_plot_site %>% 
pivot_longer(cols = -sitename, names_to = "group") %>% 
ggplot(aes(x = group, y = value, fill = as.factor(sitename))) + 
geom_bar(stat="identity", position=position_dodge())

```
The plot shows CV values for different attributes The color of the bars represent different sites. Among tehs sites, we can see SC (Sachang) site has significantly high CV values. SC has two artifacts, one is from 1st phase and the other one is 2nd phase. We assume that this time difference and few number of artifacts may relate to the high CV value.

```{r}
cv_plot_site %>%
  pivot_longer(cols = -sitename, 
               names_to = "group") %>%
  ggplot(aes(x = sitename, 
             y = value, 
             fill = as.factor(group))) +
  geom_bar(stat = "identity", 
           position = position_dodge())

```
Unlike the previous plot, in this one, the bars represent different attributes. You can see that SC has highest CV values of most attributes. Overall, SW values are very high.

Make a table of CVs for all variable grouped by phase

BM: discuss this one first, this is the overview picture for our story:

```{r}
cv_by_stage_df  <- 
df_sitename %>% 
  select(-sitename) %>% 
  group_by(SPstage.Stage) %>% 
  add_tally() %>% 
  filter(SPstage.Stage > 1) %>% 
  nest(-SPstage.Stage) %>% 
  mutate(cv_by_stage = map(data, ~map_df(.x, cv)))

cv_by_stage_df_unnest <-   
cv_by_stage_df %>% 
  unnest(cv_by_stage)

cv_plot_stage <- cv_by_stage_df_unnest %>% 
  select(-data, -n)

library(tidyr)
cv_plot_stage %>% 
pivot_longer(cols = -SPstage.Stage, names_to = "group") %>%
ggplot(aes(x = group, y = value, fill = as.factor(SPstage.Stage))) + 
geom_bar(stat="identity", position=position_dodge()) + labs(x="Attributes", fill= "Phase")
```

This plot shows the difference of CV in between 2nd and 3rd phase. For TL, both two phases have almost same values. For MW and TW, 3rd phase has lower values than the 2nd. The other 4 attributes show higher valeues in 3rd phase, which means less standardized in their shape during the latter period.

By the way, Previous research (Bettinger&Eerkins, Garvey) focused on Width of the projectile points along with weight to understand cultural transmssion. We haven't figured out which attibutes would be more infortant than others yet.

BM: this discuss this one second, this is where we get into the details

```{r}
## CV with site (n=X)

cv_by_site_df_unnest_label %>% 
  select(-SPstage.Stage) %>% 
  group_by(sitename) %>% 
  add_tally() %>% 
  filter( n > 1) %>% 
   mutate(label = as.factor(paste0(sitename,' (N = ', n, ")"))) %>%   select(-n) %>% 
 nest(-sitename, -label) %>% 
  mutate(cv_by_site = map(data, ~map_df(.x, cv)))

cv_by_site_df_unnest_label <-   
cv_by_site_df_label %>% 
  unnest(cv_by_site)

cv_plot_site_label <- cv_by_site_df_unnest_label %>%
                select(-data)

library(tidyr)
p<- cv_plot_site_label %>% 
pivot_longer(cols = -c(sitename,label), names_to = "group") %>% 
ggplot(aes(label, value, fill = as.factor(group))) + 
geom_bar(stat="identity", position=position_dodge())

p + labs(x= "Site", fill = "Attributes")

# plot CVs by site with number of artefacts showing >10
# phase 2 and phase 3 sites only
cv_by_site_df_label_more_than_10 <- 
df_sitename %>% 
  unite(col = "site_phase", 
        c(sitename, 
          SPstage.Stage), 
        sep = "_") %>% 
  group_by(site_phase) %>% 
  add_tally() %>% 
  filter( n > 10) %>% 
   mutate(label = as.factor(paste0(site_phase,' (N = ', n, ")"))) %>%   select(-n) %>% 
 nest(-site_phase, -label) %>% 
  mutate(cv_by_site = map(data, ~map_df(.x, cv))) %>% 
  unnest(cv_by_site) %>% 
  select(-data)

cv_by_site_df_label_more_than_10 %>%
  pivot_longer(cols = -c(site_phase, label), 
               names_to = "lithic_attribute",
               values_to = "CV") %>%
  ggplot(aes(x = lithic_attribute, 
             y = CV)) +
  geom_col() +
  facet_wrap(~  label) +
  ggtitle("CVs of point attributes by site (with sites >10 artefacts")

cv_by_site_df_label_more_than_10 %>%
  pivot_longer(cols = -c(site_phase, label), 
               names_to = "lithic_attribute",
               values_to = "CV") %>%
  filter(lithic_attribute %in% c("ML", "MW", "TW")) %>%

  ggplot(aes(x = lithic_attribute, 
             y = CV)) +
  geom_col() +
  facet_wrap(~  label) +
  ggtitle("CVs of point attributes by site (with sites >10 artefacts)") +
  theme_minimal()
```

Computing Pearson's correlation 
```{r}

#simpler way to calculate Correlataion, you can use
SCOR <- cor(DF)

#a scatterplot matrix of Correlation
ALL<-pairs(DF)

# correlation per phase
cor2 <- 
df_sitename %>% 
  filter(SPstage.Stage == 2) %>% 
  select(-sitename, -SPstage.Stage) %>% 
  cor(.)

ACOR2 <- pairs(cor2)

cor3 <- 
df_sitename %>% 
  filter(SPstage.Stage == 3) %>% 
  select(-sitename, -SPstage.Stage) %>% 
  cor(.)

ACOR3 <- pairs(cor3)

# Plot for SCOR
library(corrplot)

par(mfrow=c(1,3))
# correlation for all
cplotall <- corrplot(SCOR, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45, main = "Correlatoin for ALL")
# for phase 2
cplot2 <- corrplot(cor2, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45, main = "Correlatoin for Phase 2")
# for phase 3
cplot3 <- corrplot(cor3, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45, main = "Correlatoin for Phase 3")

```
=> We may need to delete this chunk since we made new plots below using corrr package and same dataset.

```{r}
#using corrr package
library(tidyverse)
library(corrr)

corrr_all <- correlate(DF) %>% 
  shave()

call<-rplot(corrr_all) +
  ggtitle("Correlation for all phase")

correlate(DF) %>% 
  network_plot()

#phase 2

DF2<-df_sitename %>% 
  filter(SPstage.Stage == 2) %>% 
  select(-sitename, -SPstage.Stage)

corrr_2 <-correlate(DF2) %>% 
  shave()

ctwo <- rplot(corrr_2) +
  ggtitle("Correlation for 2nd phase")

correlate(DF2) %>% 
  network_plot()

#phase 3

DF3<-df_sitename %>% 
  filter(SPstage.Stage == 3) %>% 
  select(-sitename, -SPstage.Stage)

corrr_3 <-correlate(DF3) %>% 
  shave()

cthree <- rplot(corrr_3) +
  ggtitle("Correlation for 3rd phase")

correlate(DF3) %>% 
  network_plot()

### three phases at once
library(patchwork)
call + ctwo + cthree + plot_layout(ncol=1)
```
When it comes to the correlations between seven attributes, BL-ML, TW-MW, TL-TW, and TL-SL are relatively close to each other. There is no negative correlation among the attributes. Compared to the 2nd phase, in 3rd phase, correlations of SL with TW and MW, and TL with TWm SL, and MW get stronger. On the other hand, the correlation of TW-SW becomes weeker. Overall, the correlation between attributes became stronger from 2nd to 3rd phase.

Combined the two results of CV and correlation from 2nd to 3rd...

CV: lower in MW and TW, but higher in other 5 attribtues.

COR: overall correlation became stronger.

=> This combination doesn't fit into the models.

Principal component analysis

Principal Component Analysis is a technique to extract relevant information from a multivariate dataset and to express this information as a set of few new variables called principal components or dimensions [@cascalheira2018use].

In archaeology, PCA has been often used for~

We applied PCA to our dataset to identify correlations between variables that would represent the existence of patterns in the points, which could possilby be translated into more ~ , in addition to correlation analysis.

```{r}
library(ggbiplot)
library(ggfortify)

PCA <- prcomp(DF, center = TRUE, scale = TRUE)

# remove stage 1 artefact
xx_pca <- 
df_sitename %>% 
  filter(SPstage.Stage != 1) %>% 
  select(-sitename, -SPstage.Stage) %>% 
  prcomp(., center = TRUE, scale = TRUE)

xx_pca1 <- 
  data.frame(xx_pca$x,
             stage = as.factor( df_sitename$SPstage.Stage[df_sitename$SPstage.Stage !=1]))

## CV with site (n=X)
label_phase <-
   df_sitename %>% 
   group_by(SPstage.Stage) %>% 
   add_tally() %>% 
   mutate(label = as.factor(paste0(SPstage.Stage,'\nN = ',n)))

count_phase <- length(which(label_phase[9]>0))
count_phase2 <- length(which(label_phase[9] == 2))
count_phase3 <- length(which(label_phase[9] == 3))

###

#plot for both stage
gg_stage <- 
ggplot(xx_pca1 ,
       aes(PC1,
           PC2,
       colour = stage)) +
  geom_point() +
  stat_ellipse()

# phase 2

xx_pca2 <- 
df_sitename %>% 
  filter(SPstage.Stage == 2) %>% 
  select(-sitename, -SPstage.Stage) %>% 
  prcomp(., center = TRUE, scale = TRUE)

xx_pca1_2 <- 
  data.frame(xx_pca2$x,
             stage = as.factor( df_sitename$SPstage.Stage[df_sitename$SPstage.Stage ==2]))

ggplot(xx_pca1_2 ,
       aes(PC1,
           PC2,
       colour = stage)) +
  geom_point() +
  stat_ellipse()

# phase 3

xx_pca3 <- 
df_sitename %>% 
  filter(SPstage.Stage == 3) %>% 
  select(-sitename, -SPstage.Stage) %>% 
  prcomp(., center = TRUE, scale = TRUE)

xx_pca1_3 <- 
  data.frame(xx_pca3$x,
             stage = as.factor( df_sitename$SPstage.Stage[df_sitename$SPstage.Stage ==3]))

ggplot(xx_pca1_3 ,
       aes(PC1,
           PC2,
       colour = stage)) +
  geom_point() +
  stat_ellipse()

```

Contribution of variables for each of four PCA dimensions of ALL phases except for the first. 
```{r}
library(gridExtra)
library(factoextra)
PC1_v <- fviz_contrib(xx_pca, choice = "var", axes = 1, top = 7)
PC2_v <- fviz_contrib(xx_pca, choice = "var", axes = 2, top = 7)
PC3_v <- fviz_contrib(xx_pca, choice = "var", axes = 3, top = 7)
PC4_v <- fviz_contrib(xx_pca, choice = "var", axes = 4, top = 7)

grid.arrange(PC1_v, PC2_v, PC3_v, PC4_v, nrow =2)

```

We see some differences in artefact shape and size. Now we need to examins the PCA output numbers to describe in more details.

```{r}

library(factoextra) #http://www.sthda.com/english/articles/31-principal-component-methods-in-r-practical-guide/112-pca-principal-component-analysis-essentials/

# Color variables by the continuous variable
cv_all <- fviz_pca_var(xx_pca, col.var = "contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07")) + ggtitle(paste("All", "(n =", count_phase,")"))

# phase 2
cv_two <- fviz_pca_var(xx_pca2, col.var = "contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07")) + ggtitle(paste("Phase 2", "(n =", count_phase2,")"))
# phase 3
cv_three <- fviz_pca_var(xx_pca3, col.var = "contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07")) + ggtitle(paste("Phase 3", "(n =", count_phase3,")"))

library(patchwork)

cv_all + cv_two + cv_three + plot_layout(ncol=1)

```
Overall, BL and ML have stronger influence than the other attributes and are related to each other but not with the other attributes throughout time. From the phase 2 to the phase 3, the proportion of TL and SL gets bigger and they are more related to SW.

SYG6 site data and analysis: artifacts from SYG6 site: now it has two layers, but the same phase (2nd)

```{r}

library(stringr)
SYG6_df <- 
  df %>% 
  filter(str_detect(files, "SYG6"))

SYG6_df_sitename <- SYG6_df %>% 
  mutate (files = str_remove_all(files, "_.tps")) %>% 
  separate (files, into = c("sitename", "files"), sep = "_") %>% 
  select (-files)

#list of all attributes without ID
SYG6_DF <- SYG6_df[,1:7]

# make n=element number 
count_SYG6 <- length(which(SYG6_df$SPstage.Stage == 2))

#simpler way to calculate Correlataion, you can use
SYG6_cor <- cor(SYG6_DF)

#a scatterplot matrix of Correlation
pairs(SYG6_DF)

cplot_SYG6 <- corrplot(SYG6_cor, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)

# using corrr

corrr_SYG6_DF <-correlate(SYG6_DF) %>% 
  shave()

cSYG6_DF <- rplot(corrr_SYG6_DF) +
  ggtitle(paste("Correlation SYG6: 2nd Phase", "(n =", count_SYG6,")"))

correlate(SYG6_DF) %>% 
  network_plot()

# CV
map_df(SYG6_DF, cv)

SYG6_PCA <- prcomp(SYG6_cor, center = TRUE, scale = TRUE)

# remove stage 1 artefact
SYG6_xx_pca <- 
SYG6_df_sitename %>% 
  select(-sitename, -SPstage.Stage) %>% 
  prcomp(., center = TRUE, scale = TRUE)

SYG6_xx_pca1 <- 
  data.frame(SYG6_xx_pca$x,
             stage = as.factor( SYG6_df_sitename$SPstage.Stage))

#plot for both stage
SYG6_gg_stage <- 
ggplot(SYG6_xx_pca1 ,
       aes(PC1,
           PC2,
       colour = stage)) +
  geom_point() +
  stat_ellipse()

cv_SYG6 <-fviz_pca_var(SYG6_xx_pca, col.var = "contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07")) +
             ggtitle(paste("SYG6: 2nd Phase", "(n =", count_SYG6,")"))

SYG6_PC1_v <- fviz_contrib(SYG6_xx_pca, choice = "var", axes = 1, top = 7)
SYG6_PC2_v <- fviz_contrib(SYG6_xx_pca, choice = "var", axes = 2, top = 7)

grid.arrange(SYG6_PC1_v, SYG6_PC2_v, nrow =1)

```

SYG1 site data and analysis: artifacts from SYG1 site, which include two differnt locations(excavation pits) and the artifacts are from different phase.
SYG1_2 : 3rd phase
SYG1_1: 2nd phase

```{r}
#SYG1_2
library(stringr)
SYG1_2df <- 
  df %>% 
  filter(str_detect(files, "SYG1_2"))

library(dplyr)
SYG1_2df_sitename <- SYG1_2df %>% 
  mutate (files = str_remove_all(files, "_\\d+.tps")) %>% 
  dplyr::rename(sitename = files)

#list of all attributes without ID
SYG1_2DF <- SYG1_2df[,1:7]

# make n=element number 
count_SYG1_2 <- length(which(SYG1_2df$SPstage.Stage == 3))

#simpler way to calculate Correlataion, you can use
SYG1_2cor <- cor(SYG1_2DF)

#a scatterplot matrix of Correlation
pairs(SYG1_2DF)

cplot_SYG1_2 <- corrplot(SYG1_2cor, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)

# using corrr

corrr_SYG1_2DF <-correlate(SYG1_2DF) %>% 
  shave()

cSYG1_2DF <- rplot(corrr_SYG1_2DF) +
  ggtitle(paste("Correlation SYG1_2: 3rd Phase", "(n =", count_SYG1_2,")"))

correlate(SYG1_2DF) %>% 
  network_plot()

# CV
map_df(SYG1_2DF, cv)

SYG1_2PCA <- prcomp(SYG1_2DF, center = TRUE, scale = TRUE)

# remove stage 1 artefact
SYG1_2xx_pca <- 
SYG1_2df_sitename %>% 
  select(-sitename, -SPstage.Stage) %>% 
  prcomp(., center = TRUE, scale = TRUE)

SYG1_2xx_pca1 <- 
  data.frame(SYG1_2xx_pca$x,
             stage = as.factor(SYG1_2df_sitename$SPstage.Stage))

#plot for both stage
SYG1_2gg_stage <- 
ggplot(SYG1_2xx_pca1 ,
       aes(PC1,
           PC2,
       colour = stage)) +
  geom_point() +
  stat_ellipse()

cv_SYG1_2 <- fviz_pca_var(SYG1_2xx_pca, col.var = "contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07")) +
  ggtitle(paste("SYG1_2: 3rd Phase", "(n =", count_SYG1_2,")"))

SYG1_2PC1_v <- fviz_contrib(SYG1_2xx_pca, choice = "var", axes = 1, top = 7)
SYG1_2PC2_v <- fviz_contrib(SYG1_2xx_pca, choice = "var", axes = 2, top = 7)

grid.arrange(SYG1_2PC1_v, SYG1_2PC2_v, nrow =1)

```

```{r}
#SYG1_1
library(stringr)
SYG1_1df <- 
  df %>% 
  filter(str_detect(files, "SYG1_1"))

library(dplyr)
SYG1_1df_sitename <- SYG1_1df %>% 
  mutate (files = str_remove_all(files, "_\\d+.tps")) %>% 
  dplyr::rename(sitename = files)

#list of all attributes without ID
SYG1_1DF <- SYG1_1df[,1:7]

# make n=element number 
count_SYG1_1 <- length(which(SYG1_1df$SPstage.Stage == 2))

#simpler way to calculate Correlataion, you can use
SYG1_1cor <- cor(SYG1_1DF)

#a scatterplot matrix of Correlation
pairs(SYG1_1DF)

cplot_SYG1_1 <- corrplot(SYG1_1cor, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)

# using corrr

corrr_SYG1_1DF <-correlate(SYG1_1DF) %>% 
  shave()

cSYG1_1DF <- rplot(corrr_SYG1_1DF) +
  ggtitle(paste("Correlation SYG1_1: 2nd Phase", "(n =", count_SYG1_1,")"))

correlate(SYG1_1DF) %>% 
  network_plot()

# CV
map_df(SYG1_1DF, cv)

SYG1_1PCA <- prcomp(SYG1_1DF, center = TRUE, scale = TRUE)

# remove stage 1 artefact
SYG1_1xx_pca <- 
  SYG1_1df_sitename %>% 
  select(-sitename, -SPstage.Stage) %>% 
  prcomp(., center = TRUE, scale = TRUE)

SYG1_1xx_pca1 <- 
  data.frame(SYG1_1xx_pca$x,
             stage = as.factor(SYG1_1df_sitename$SPstage.Stage))

#plot for both stage
SYG1_1gg_stage <- 
ggplot(SYG1_1xx_pca1 ,
       aes(PC1,
           PC2,
       colour = stage)) +
  geom_point() +
  stat_ellipse()

cv_SYG1_1 <- fviz_pca_var(SYG1_1xx_pca, col.var = "contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07")) +
            ggtitle(paste("SYG1_1: 2nd Phase", "(n =", count_SYG1_1,")"))

SYG1_1PC1_v <- fviz_contrib(SYG1_1xx_pca, choice = "var", axes = 1, top = 7)
SYG1_1PC2_v <- fviz_contrib(SYG1_1xx_pca, choice = "var", axes = 2, top = 7)

grid.arrange(SYG1_1PC1_v, SYG1_1PC2_v, nrow =1)

```

Correlations for each site
```{r}
par(mfrow=c(1,3))
cplot_SYG6_ <- corrplot(SYG6_cor, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45, title = "SYG6")
cplot_SYG1_1 <- corrplot(SYG1_1cor, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45, title = "SYG1_1")
cplot_SYG1_2 <- corrplot(SYG1_2cor, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45, title = "SYG1_2")

### three sites at once, using corrr
library(patchwork)
cSYG6_DF + cSYG1_2DF + cSYG1_1DF + plot_layout(ncol=1)

```
In this plot, we can see that the pattern of correlation for SYG6 and SYG1_1 are similar in terms of correlation based on ML/SL/TL with other attributes. Both are part of the phase 2. However, SYG1_2 is younger than the other two and the correlation based on SW become weaker. 
When it comes to the difference between SYG6 and SYG1_1, the correlation between artifacts of SYG1_1 is stronger than SYG6.

PCA for each site
```{r}
cv_SYG6 + cv_SYG1_1 + cv_SYG1_2 +  plot_layout(ncol=1)
```
Unlike the correlation of the three units, PCA of the three are more various. Especially, SW is related to SL/TL in SYG6, related to ML/BL in SYG1_1, and unrelated to the other attributes in SYG1_2.

```{r}
#I'm working on how to read oulines and how to use them for anlaysis. We may not use this for the research.
# reading landmarks and outline in tps, code from Morpho package, adjust readallTPS function.

#readalltps = function(file) 
{
   # out = list()
   # exOut <- FALSE
  # noutline <- 0
  # input <- readLines(file)
   # LM <- grep("LM=", input)
  # LMstring <- input[LM]
   # nLM <- sapply(LMstring, strsplit, split = "=")
  #  nLM <- unlist(nLM)
   # nLM <- as.integer(nLM[-which(nLM == "LM")])
  #  nobs <- length(nLM)
   # ID <- grep("ID=", input)
  #  IDstring <- input[ID]
   # nID <- sapply(IDstring, function(x) {
   #     x <- gsub("=", "_", x)
  #  })
   # nID <- gsub(" ", "", unlist(nID))
  #  out$ID <- nID
   # outline <- grep("CURVES=", input)
  #  if (length(outline) > 0) {
   #     exOut <- TRUE
    #    outlinestring <- input[outline]
     #   noutline <- sapply(outlinestring, strsplit, split = "=")
      #  noutline <- unlist(noutline)
      #  noutline <- as.integer(noutline[-which(noutline == "CURVES")])
    #    LMoutline <- (sapply(outline, function(x) {
   #         x <- max(which(LM < x))  })) }
 #   LMdata <- list()
  #  for (i in 1:nobs) {
  #      if (nLM[i] > 0) {
   #         LMdata[[i]] <- as.numeric(unlist(strsplit(unlist(input[c((LM[i] + 
      #          1):(LM[i] + nLM[i]))]), split = " ")))
       #     LMdata[[i]] <- matrix(LMdata[[i]], nLM[i], 2, byrow = TRUE)
        }
   #     else LMdata[[i]] <- NA }
   # names(LMdata) <- nID
  #  out$LM <- LMdata
   # if (exOut) {
    #    outlineData <- list()
     #   for (i in 1:nobs) {
      #      if (i %in% LMoutline) {
       #         i1 <- grep(i, LMoutline)
        #        outlinetmp <- list()
         #       ptr <- outline[i1] + 1
          #      j <- 1
           #     while (j <= noutline[i1]) {
            #      tmpnr <- as.integer(unlist(strsplit(input[ptr], 
        #            split = "="))[2])
         #         outlinetmp[[j]] <- matrix(as.numeric(unlist(strsplit(input[(ptr + 
         #           1):(ptr + tmpnr)], split = " "))), tmpnr, 
          #          2, byrow = TRUE)
           #       ptr <- ptr + tmpnr + 1
            #      j <- j + 1  }
             #   outlineData[[i]] <- outlinetmp  }
          #  else outlineData[[i]] <- NA }
       # names(outlineData) <- nID
      #  out$outlines <- outlineData }
 #   cat(paste("Read", nobs, "datasets with", sum(noutline), "outlines\n"))
#    return(out)}

#try to read all data(landmarks and outlines) from one file
# lm_trial <- readalltps(here("analysis/data/raw_data/A_BG_2.tps"))

#read all files from raw_data
# tps_files <- list.files(here("analysis/data/raw_data"), full.names = TRUE)
# lm_landmarks <- map(tps_files, readalltps)
# names(lm_landmarks) <- list.files(here("analysis/data/raw_data"))

#lm_landmarks = c()
#for (i in 1:length(files)) {
#  filename <- paste(datadir, files[i], sep = "")
 # print(filename)
#  landmarks <- readalltps(filename)
# lm_landmarks <- c(lm_landmarks, landmarks)}

```

# Results

```{r get-data, eval = FALSE}
# Note the path that we need to use to access our data files when rendering this document
my_data <- read.csv(here::here('analysis', 'data', 'raw_data', 'my_csv_file.csv'))
```

# Discussion

# Conclusion

# Acknowledgements

<!-- The following line inserts a page break when the output is MS Word. For page breaks in PDF, use \newpage on its own line.  -->
##### pagebreak

# References 
<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
<div id="refs"></div>

##### pagebreak

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
git2r::repository(here::here())
```