### 2021-05-12 analysis/paper/garvey_ct_simulation.Rmd

---
title: "CT simulation"
author: "R Garvey"
date: "6/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read data
```{r}
library(here)
data <- read.csv(here("analysis/data/garvey_data.csv"))
str(data)
```

### Multi panel Washita figure

```{r}

# Plot and simulation parameters

#######################

D <- data.frame(data$wash_width, data$wash_basewidth,
                data$wash_neck, data$wash_haftlength,
                data$wash_length, data$wash_midlength, 
                data$wash_thick, data$wash_bladelength, 
                data$wash_weight)

# set counter to # variables (i.e., # figure rows)
count <- c(1:9)

# set the attribute-specific n.measures
n.measures <- c(162, 157,
                249, 227,
                136, 136,
                259, 135,
                87)

# y-axis panel plot labels
labs <- c("width\n(mm)", "base width\n(mm)",
          "neck width\n(mm)", "haft length\n(mm)",
          "length\n(mm)", "mid length\n(mm)",  
          "thickness\n(mm)", "blade length\n(mm)",
          "weight\n(g)")

## simulation parameters
set.seed(1212)            # set seed
n.houses       <- 100     # set number of houses at site
y              <- 1000    # set number of simulation runs

## plotting parameters
col.density   <- "gray80"   # set density color
col.outline   <- "gray20"   # set density outline color
col.mean      <- "black"    # set col for mean of arch sample
cex.ylab      <- 1.2        # set y-axis label size
cex.ylabtick  <- 1.1        # set y-axis tick label size
cex.xlabtick  <- 1.1        # set x-axis tick label size

x.lim2   <- c(6, 7,         # set x limit for each plot
              3.5, 4,       # (this also affects tick mark placement)
              50, 40,
              1.4, 35,
              0.12)

y.lim2   <- c(1.2, 1,       # set y limit for each plot
              1.75, 1.2,    
              0.15, 0.15,
              5, 0.15,
              45)

###############################

# The plot:

pdf(file="/Users/##ENTER DESTINATION##/Garvey_Figure3.pdf", 
    height = 12, width = 12)

par(mfcol = c(length(count), 3))
par(oma = c (6, 10, 5, 2))

# column 1 (SD = 0.1)

for(x in count) {   # looping procedure for each plotting pannel

SD       <- 0.1    # set standard deviation multiplier for random draws 
d        <- na.omit(D[, x])     # select the data set to sample
N        <- n.measures[x]       # select the appropriate n.measures

metadig   <- NULL  #create empty vector to hold digs
var.Kdigs <- NULL  #create empty vector to hold variances
cv.Kdigs  <- NULL  #create empty vector to hold CVs

for(i in 1:y) {    # "master" loop that  repeats
                   # "archaeological sampling"" from 
                   # "generational learning sample"" one thousand times

    all.houses <-  NULL         # create an empty vector to hold results

    for(j in 1:n.houses) {       #loop through all houses

               house.seed <- sample(d, 1, replace = TRUE)   
               house.g1 <- rnorm(4, mean = house.seed, 
                                 sd = SD * mean(house.seed)) 
               house.g2 <- rnorm(4, mean = mean(house.g1), 
                                 sd = SD * mean(house.g1)) 
               house.g3 <- rnorm(4, mean = mean(house.g2), 
                                 sd = SD * mean(house.g2))
               house.g4 <- rnorm(4, mean = mean(house.g3), 
                                 sd = SD * mean(house.g3))
               total.house.sample <- as.vector(c(house.g1,
                                                  house.g2,
                                                  house.g3,
                                                  house.g4))
           all.houses <- c(total.house.sample, all.houses)

        }
           dig <- sample(all.houses, N, replace = TRUE)
           metadig <- c(metadig, dig)
           var.Kdigs <- c(var(dig), var.Kdigs)
           cv.Kdigs <- c(sd(dig)/mean(dig), cv.Kdigs)
}

#plotting
dp.md <- density(var.Kdigs)
par(mai = c(0.25, 0.5, 0.1, 0))

plot(dp.md, col = col.outline, bty = 'n',
     main = "", xlab = "", ylab = "",
     xlim = c (0, x.lim2[x]),
     ylim = c (0, y.lim2[x]),
     xaxt = "n", yaxt = "n")

axis(1, at = c(0, 0.25 * x.lim2[x],
               0.5 * x.lim2[x], 0.75 * x.lim2[x],
               x.lim2[x]),
     cex = cex.xlabtick)

# alternative, with only 3 x-axis ticks
# axis(1, at = c(0, 0.5 * x.lim2[x],
#                x.lim2[x]),
#      cex = cex.xlabtick)

axis(2, at = c(0, 0.5 * y.lim2[x],
               y.lim2[x]),
     line = 1.5, las = 2,
     cex = cex.ylabtick)

polygon(dp.md, col = col.density)
abline(v = var(d), col = col.mean, lwd = 2)
mtext (text = labs[x], side = 2, 
       las = 1, line = 5.5,
       cex = 1.2, outer = FALSE)

}

mtext (text = "A. CV = 10%", side = 3, 
       las = 1, line = 73,
       cex = 1.5, outer = FALSE)

# column 2 (CV = 5%)

for(x in count) {   # looping procedure for each plotting pannel

SD       <- 0.05    # set standard deviation multiplier for random draws 
d        <- na.omit(D[, x])     # select the data set to sample
N        <- n.measures[x]       # select the appropriate n.measures

metadig   <- NULL  #create empty vector to hold digs
var.Kdigs <- NULL  #create empty vector to hold variances
cv.Kdigs  <- NULL  #create empty vector to hold CVs

for(i in 1:y) {    # "master" loop that should repeat
                   # "archaeological sampling"" from 
                   # "generational learning sample"" one thousand times

    all.houses <-  NULL         # create an empty vector to hold results

    for(j in 1:n.houses) {       #loop through all houses

               house.seed <- sample(d, 1, replace = TRUE)   
               house.g1 <- rnorm(4, mean = house.seed, 
                                 sd = SD * mean(house.seed)) 
               house.g2 <- rnorm(4, mean = mean(house.g1), 
                                 sd = SD * mean(house.g1)) 
               house.g3 <- rnorm(4, mean = mean(house.g2), 
                                 sd = SD * mean(house.g2))
               house.g4 <- rnorm(4, mean = mean(house.g3), 
                                 sd = SD * mean(house.g3))
               total.house.sample <- as.vector(c(house.g1,
                                                  house.g2,
                                                  house.g3,
                                                  house.g4))
           all.houses <- c(total.house.sample, all.houses)

        }
           dig <- sample(all.houses, N, replace = TRUE)
           metadig <- c(metadig, dig)
           var.Kdigs <- c(var(dig), var.Kdigs)
           cv.Kdigs <- c(sd(dig)/mean(dig), cv.Kdigs)
}

#plotting
dp.md <- density(var.Kdigs)
par(mai = c(0.25, 0.5, 0.1, 0))

plot(dp.md, col = col.outline, bty = 'n',
     main = "", xlab = "", ylab = "",
     xlim = c (0, x.lim2[x]),
     ylim = c (0, y.lim2[x]),
     xaxt = "n", yaxt = "n")

axis(1, at = c(0, 0.25 * x.lim2[x],
               0.5 * x.lim2[x], 0.75 * x.lim2[x],
               x.lim2[x]),
     cex = cex.xlabtick)

polygon(dp.md, col = col.density)
abline(v = var(d), col = col.mean, lwd = 2)

}

mtext (text = "B. CV = 5%", side = 3, 
       las = 1, line = 73,
       cex = 1.5, outer = FALSE)

# column 3 (CV = 3%)

for(x in count) {   # looping procedure for each plotting pannel

SD       <- 0.03    # set standard deviation multiplier for random draws 
d        <- na.omit(D[, x])     # select the data set to sample
N        <- n.measures[x]       # select the appropriate n.measures

metadig   <- NULL  #create empty vector to hold digs
var.Kdigs <- NULL  #create empty vector to hold variances
cv.Kdigs  <- NULL  #create empty vector to hold CVs

for(i in 1:y) {    # "master" loop that should repeat
                   # "archaeological sampling"" from 
                   # "generational learning sample"" one thousand times

    all.houses <-  NULL         # create an empty vector to hold results

    for(j in 1:n.houses) {       #loop through all houses

               house.seed <- sample(d, 1, replace = TRUE)   
               house.g1 <- rnorm(4, mean = house.seed, 
                                 sd = SD * mean(house.seed)) 
               house.g2 <- rnorm(4, mean = mean(house.g1), 
                                 sd = SD * mean(house.g1)) 
               house.g3 <- rnorm(4, mean = mean(house.g2), 
                                 sd = SD * mean(house.g2))
               house.g4 <- rnorm(4, mean = mean(house.g3), 
                                 sd = SD * mean(house.g3))
               total.house.sample <- as.vector(c(house.g1,
                                                  house.g2,
                                                  house.g3,
                                                  house.g4))
           all.houses <- c(total.house.sample, all.houses)

        }
           dig <- sample(all.houses, N, replace = TRUE)
           metadig <- c(metadig, dig)
           var.Kdigs <- c(var(dig), var.Kdigs)
           cv.Kdigs <- c(sd(dig)/mean(dig), cv.Kdigs)
}

#plotting
dp.md <- density(var.Kdigs)
par(mai = c(0.25, 0.5, 0.1, 0))

plot(dp.md, col = col.outline, bty = 'n',
     main = "", xlab = "", ylab = "",
     xlim = c (0, x.lim2[x]),
     ylim = c (0, y.lim2[x]),
     xaxt = "n", yaxt = "n")

axis(1, at = c(0, 0.25 * x.lim2[x],
               0.5 * x.lim2[x], 0.75 * x.lim2[x],
               x.lim2[x]),
     cex = cex.xlabtick)

polygon(dp.md, col = col.density)
abline(v = var(d), col = col.mean, lwd = 2)

}

mtext (text = "C. CV = 3%", side = 3, 
       las = 1, line = 73,
       cex = 1.5, outer = FALSE)

dev.off()
```