### 2023-02-14 analysis/paper/paper.qmd
commit to merge with Ben's edits

---
title: "Cultural Transmission and Technological Transitions during the Late Paleolithic in Korea"
author:
  - Gayoung Park:
      institute: uw
      correspondence: "yes"
      email: gayoungp@uw.edu
  - Ben Marwick:
      institute: uw
institute:
  - uw:
      name: Department of Anthropology, University of Washington
      address: Seattle, WA, USA
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: false
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
  cache: false
editor: visual
date: now
date-format: long
title-block-published: "Last updated" 
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style

abstract: |
  The onset of the Late Paleolithic period in Korea, represented by the appearance of stemmed points and blades, was a key event in the dispersal of modern humans in Northeast Asia. Previous studies have mainly focussed on possible origin locations and routes of these new technologies. The specific cultural processes of the appearance stemmed points and blades has rarely been considered. In this research, we investigate the potential social contexts related to the technological transition by applying a cultural transmission framework. Our main question is: what was the dominant mode of cultural transmission during this time of technological innovation in the Korean Late Paleolithic? Following Bettinger and Eerkens [-@bettinger1999point], we tested models of guided variation and indirect bias using data from the Korean Paleolithic assemblages containing stemmed points. We computed correlation coefficients and coefficients of variation (CV) to evaluate the transmission biases. Based on our results, it appears that information about the novel technology was transmitted via guided variation with a low impact of indirect bias. Some attributes, such as length and width, appear to be transmitted with little variation than others. We assume that individuals or groups developed stemmed points by experimenting with blade technologies. And then they copied the key elements of a successful model to ensure quality and to optimize tool performance. The shape of stemmed points became more standardized among their social groups as a result.

keywords: |
  Stemmed points; Korean Paleolithic; Cultural Transmission; Learning process, Copying errors, Transmission Biases, Guided variation, Indirect bias
---

# Introduction {.unnumbered}

The application of evolutionary theory to archaeological research has been favored for the study of technological transitions and related human behaviors in the remote past [@dunnell1980evolutionary; @lipo1997population; @bettinger1997rediscovering; @bettinger1999point; @mesoudi2008cultural].  Archaeologists have used evolutionary theories and methods to study human behavioral ecology, cultural transmission, and artifact phylogenetics in the past [@riede2010isn; @o2017dual; @garvey2018current]. In this paper we use cultural transmission theory to investigate technological transitions during the Korean Late Paleolithic. The primary technological innovation of this period was the introduction of stemmed points and blades. These new lithic technologies may represent the first arrival of modern humans in this region, and thus they may also mark an important event in human dispersal through East Asia [@seong2009emergence]. Most previous studies on stemmed points focused on where they may have originated, connecting Korea with global patterns of modern human dispersal. There are, however, largely unanswered questions about the specific cultural processes and social contexts of this technological change in Korea [@seong2008tanged; @bae2010origin; @bae2017origin]. Our study explores the social contexts in which new technologies emerged in the Korean Late Paleolithic based on a cultural transmission framework. Our main question is: What was the dominant mode of cultural transmission for technological innovation in the Korean Late Paleolithic? We also ask: Do modes of cultural transmission vary over time and space? We consider three possible modes of cultural transmission: guided variation (trial and error), indirect bias (copying others), or a combination of the two. The results of this study have implications for determining whether these novel technologies originated outside of the Korean Peninsula or if they were locally developed independently.

# The Late Paleolithic of the Korean Peninsula {.unnumbered}

The emergence of stemmed points marks the beginning of the Late Paleolithic period in Korea around 40-35 ka. A stemmed point is a projectile point made out of an elongated flake or blade with a slight retouch on the proximal end to shape an acute tip, and on the distal end to make a stem, which connects to a shaft. Stemmed points were the first composite tool types to appear on the Korean Peninsula, and were a symbol of new hunting strategies there and in adjacent regions [@seong2008tanged; @lee2019were]. A number of technological innovations are evident on stemmed points, including blade technology, multiple manufacturing stages, and evidence of resharpening and reusing, which were rarely seen earlier [@chang2013human; @bamforth2009projectile; @seong2015diversity]. Currently, the oldest stemmed points in Northeast Asia are from the Yongho-doing site in Korea, dating back to 38.5ka [@seong2009emergence; @seong2015diversity; @bae2017origin]. Following their appearance in Korea, stemmed points spread to the Japanese archipelago [@chang2013human].

A majority of previous studies of the technological transitions during the Late Paleolithic in Korea have focused on the possible origin locations of stemmed points, as well as the timing and routes of modern human dispersal. In order to explain the origin of stemmed points, two competing models can be distinguished: *in situ* evolution [@seong2009emergence] and heterogenic migration [@bae2010origin]. The *in situ* model claims that stemmed points and other Late Paleolithic assemblages, including blade industries, autonomously developed on the Korean peninsula as a result of convergent evolution [@seong2006structure; @seong2008tanged; @seong2009emergence]. In order to support his claim, @seong2009emergence examined the blade-to-flake ratios of stone artifact assemblages in South Korea. In his view, the blade industry represents a novel invention, while flake materials indicate a continuously used technology. Based on the increased proportion of blades in stone artifact assemblages during the Late Paleolithic, he argues that the new technologies spread throughout the area after its local invention. In addition, Seong argues that the emergence of new technologies is a result of gradual, local, evolutionary processes, as evidenced by the increasing number of stemmed points over time.

On the other hand, the migration model argues that the new blade industry, including stemmed points, and the earlier simple flake tool tradition, including large cores, polyhedrals, choppers, and handaxes, came from different origins [@bae2010origin; @bae2012nature; @bae2013early]. While the *in situ* model claims that the heterogenic character is the result of indigenous development, the migration model proposes that it is the result of the continuous influx of modern human populations from both north and south. The blade technology was introduced from Siberia, Mongolia, or other regions of northeast China following the Liaohe and Sunghe rivers around 35 ka BP, while the simple flake tool tradition came from  southern China [@bae2013early]. The migration of people from the southern route is supported by genetic studies of the Y chromosome, indicating that the O3-M122 M122 haplotype originated from southern East Asia and moved to northeastern Asia, including Korea, at 30-25 ka BP [@shi2005chromosome]. Bae et al. [-@bae2012nature] assume this southern migration could be related to paleoenvironmental fluctuations during the MIS 3 to 2 transition, which made the Yellow Sea/West Sea area more accessible.

According to Lee [-@lee2013current], the transition from the Korean Early Paleolithic to the Korean Late Paleolithic might have been more complicated than either migration models or *in situ* development models. Under conditions of low effective population size, he partially agrees with the *in situ* model that simple flake tools had continuously used in Korea due to ancestor-descendant relationships. With regard to the blade industry, he claims that low degrees of uniformity and small quantities of blade-associated toolkits indicate an origin outside of Korea, perhaps resulting from trade or migration. We explore these three options, *in situ*, migration, and a mixture of the two, by measuring transmission biases in assemblages of stemmed points.

# Cultural transmission and transmission biases {.unnumbered}

We use Darwinian evolutionary theory to measure transmission biases in tool-making. The theory has helped archaeologists better understand human behaviors and technological innovations, for example, using human behavioral ecology, phylogenetics, cultural transmission, and niche construction [@dunnell1980evolutionary; @richerson1992cultural; @lipo1997population; @bettinger1997evolutionary; @bettinger1999point; @eerkens2008cultural; @mesoudi2008cultural]. Cultural evolutionary approaches are effective at explaining cultural changes through mechanisms of inheritance, variation-generating processes, and selection. Although cultural systems follow similar mechanisms to biological systems, they have unique characteristics. Social transmission strongly influences these key evolutionary processes [@Whiten_2017], and can occur through a variety of learning processes, including stimulus enhancement, emulation, imitation, and teaching [@o2000applying; @schillinger2014copying].  In this paper, we examine how different social contexts can result in different modes of social transmission of tool-making skills among hunter-gatherer communities.

######
Cultural transmission theory holds that information about how to behave is acquired through interaction with other individuals and the environment [@boyd1988culture; @richerson1992cultural]. Individuals learn by themselves (e.g. trial and error), or from each other by sharing information. Information can be modified (also known as 'biased') depending on an individual's transmission context and cultural repertoire. Modification of information can occur by recombination, loss, or partial alteration [@eerkens2005cultural; @o2017dual]. Transmission biases can be important loci of changes in material culture, and can be influenced by the social contexts of cultural transmission [@heyes1994social; @eerkens2007cultural; @lycett2015cultural; @creanza2017cultural; @kendal2018social]. These biases include guided variation (where individuals learn new behaviors through trial and error), content-based bias (where some aspect of the transmitted instructions, such as cultural preferences, makes them more likely to be adopted), frequency-based bias (where an individual is biased to choose particular behaviors based on their perceived frequency in the population, such as extremely popular or rare behaviors), and indirect bias (where a behavior is transmitted because of its association with other attributes, such as the prestige or skill of other individuals) [@boyd1988culture; @richerson1992cultural; @o2017dual].

Bettinger and Eerkens [-@bettinger1997evolutionary; -@bettinger1999point; -@eerkens2008cultural] have shown how two of these biases can be identified in stone artifact assemblages. Their research focused on metric variables of stone points during the introduction of bow and arrow technology in the Great Basin around AD 300-600. They equated guided transmission (where individuals learn new behaviors through trial and error) with high metric variation and low correlation between metric attributes. Conversely, they inferred indirect bias (where a behavior is transmitted because of its association with other attributes, such as the prestige or skill of other individuals) from less variation and more correlated variables. They found that artifacts in the new bow and arrow technology in eastern California have low correlations of basal width and mass, which they interpreted as a result of guided transmission dominating the introduction and spread of these tools. From these results, they inferred a social context in eastern California of distant and unfamiliar neighbors, with limited contact between groups to acquire technology directly. In this context of limited contact, they argue that the new technologies developed largely by trial and error. On the other hand, in central Nevada they found that metric variables are highly correlated, indicating that the bow and arrow was introduced and spread by indirect bias.

Following Bettinger and Eerkens' framework, Garvey [-@garvey2018current] uses simulation to explore the degree of standardization represented by coefficient of variation (CV) values of projectile points from the US Southwest and westernmost southern High Plains. Garvey measured CV values of weight, thickness, width and lengths on two types of projectile points, Washita and Fresno points, from the Henderson site. She then computed simulated CV values according to three scenarios of different levels of transmission fidelity (i.e. CV = 10%, 5%, and 3%). The observed metrics of the archaeological projectile points are closest to the metrics of the simulated assemblage with 3% CV, which represents "extremely high-fidelity copying". A CV of 3% is the Weber fraction, or threshold of human visual perception, with variation below this value being too subtle for people to notice [@eerkens2000practice]. Garvey's work shows how the cultural transmission of tool-making behaviours can be measured, simulated, and used to interpret the archaeological record.

Garvey and Bettinger and Eerkens inferred transmission processes from material culture because they were unable to observe them directly. Direct observations have been reported by Mesoudi and O'Brien [-@mesoudi2008cultural] who conducted experimental research using groups of undergraduate students to test the underlying assumptions of Bettinger and Eerkens' study. They simulated model-based bias in projectile point making by providing the design of a model and information of the model's prior success to their research participants. They also simulated guided variation by allowing their participants to explore their own designs. They observed that the majority of participants who were able to choose the previous design copied the most-successful model. Metric attributes of the points made when copied from successful models were more highly correlated than attributes of points made by trial and error. Mesoudi and O'Brien's experimental results are important because they confirm the robustness of previous assumptions about cultural transmission biases, and validate the connection between material culture variation and cultural transmission biases.

# Modeling the social context of the appearance of stemmed points in the Late Paleolithic of Korea {.unnumbered}

Inspired by Bettinger and Eerkens' approach, we use two contrasting transmission modes, guided variation and indirect bias, to investigate the spread of stemmed point technology during the Korean Late Paleolithic period. We propose a spectrum on which we can locate foraging groups and how they started to make stemmed points, depending on their degree of social isolation or social connectedness. On one end of the spectrum we have socially isolated groups who made stemmed points entirely stimulated by guided variation; and on the other end of the spectrum we have socially connected groups whose knowledge of stemmed points derived from indirect bias on the transmission process.

To contextualize this further, on one end of our spectrum we have socially isolated groups who stayed in physically remote places from other groups, or had unfamiliar neighbors with limited contact between groups. Our assumption for this social context is that individuals or groups acquired the technology of stemmed points by modifying existing flake tool forms through trial-and-error processes to solve problems relating to resource procurement (cf. Seong's *in situ* model). This trial and error behavior is expected to have left a distinctive signature on metric variables of the stemmed points. If trial and error was the dominant bias in the transmission of knowledge about how to make stemmed points, we expect that the morphological attributes on a stemmed point will be poorly correlated and show high variation within assemblages and between sites.

On the other end of our spectrum, we have socially connected groups that occupied places close to other groups or had regular contact with other groups. Relatively high degrees of social connectivity provide frequent opportunities for observing others and acquiring information. Individuals or groups learned the technology of stemmed points by copying a model from another individual or group. In this scenario, the model, or ideal, stemmed point is likely to have been a highly successful and frequently chosen one, and learners copy all information about the point design as a package (cf. Bae's migration model). In this social context, indirect bias is the dominant influence on transmission of lithic technology. If indirect bias was prominent in the Korean Paleolithic, we expect that stemmed points should be more standardized, and attributes on the stemmed points will be more correlated, and assemblages from multiple sites will show less variance between and within them.

# Materials and methods {.unnumbered}

```{r}
#| label: summary-of-data

library(tidyverse)
library(ggplot2)
library(ggbeeswarm)

# import the cleaned data ready to work on
load(here::here("analysis/data/derived_data/data_main.RData"))

# how many stemmed artifacts per site?
stemmed_points_per_site <-
df_full_sitename %>%
  dplyr::count(full_sitename) %>%
  arrange(desc(n))

# how many assemblages do we use?
stemmed_points_ass <- df_sitename %>%
  unite(col = "site_phase",
        c(sitename,
          SPstage.Stage),
        sep = "_") %>%
  dplyr::select(-SPstage.Raw_material) %>%
  dplyr::count(site_phase)

# prepare some values to use in the text below
total_number_of_sp <- sum(stemmed_points_per_site$n)
total_number_of_sites <-nrow(stemmed_points_per_site)
total_number_of_assemblages <- nrow(stemmed_points_ass)
```

## Materials and stemmed points chronology

After the first discovery of stemmed points at the Seokjangri site in the 1960s, around 300 have been found in nearly 30 sites across Korea (@fig-map). Most sites contain only a few points but a few sites have many more, such as Suyanggae (n = 55), Jingeuneul (n = 99), and Yongsandong (n = 38) [@Kim_2017]. Among these stemmed points, we selected those that are unbroken from the tip to the stem. We excluded artifacts that were recorded as stemmed points but lack a stem. This resulted in a sample of `r total_number_of_sp` stemmed points from `r total_number_of_assemblages` assemblages unearthed from `r total_number_of_sites` sites spanning the period 40-17 ka (@tbl-sp-number). The images of the stemmed points were obtained from published excavation reports and by direct photography during our research on local museum collections. We defined multiple assemblages in a site where artifact-bearing deposits were separated by culturally sterile deposits, or where distinct artifact-bearing stratigraphic units could be identified by major differences in the texture, color, and composition of the sedimentary deposits.

\newpage

```{r}
#| label: fig-map
#| fig-cap: "Korean Paleolithic sites mentioned in this study."
#| fig-width:   4.45
#| fig-height:  4.5

# This script draws the map, this is figure 1 in the paper.

korean_archaeological_site_locations <-
  readr::read_csv(here::here("analysis/data/raw_data/sitelocation.csv"))

korean_archaeological_site_locations <-
  tibble::rowid_to_column(korean_archaeological_site_locations, "ID")

# load map tiles we previously saved, we got these using the code in 
# 000-download-map-tiles.R

library(ggrepel)
# devtools::install_github('3wen/legendMap')
library(legendMap)
library(sp)
library(ggmap)

# load map tiles we previously saved
map <- readRDS(here::here("analysis/data/raw_data/map_tiles.rds"))

# map with site names next to the points
map_of_sites <-
  ggmap(map)  +
  geom_point(data = korean_archaeological_site_locations,
             aes(long_dd ,
                 lat_dd),
             colour = "red",
             size = 2) +

# label points with site name
  geom_text_repel(data = korean_archaeological_site_locations,
                  aes(long_dd ,
                      lat_dd,
                      label = full_sitename),
                  size = 2,
                  bg.color = "white",
                  bg.r = 0.15) +
  theme_minimal(base_size = 6) +
  labs(x = "Longitute",
       y = "Latitude") +
  legendMap::scale_bar(
    # edit these numbers to select a suitable location
    # for the scale bar where it does not cover
    # important details on the map
    lon = 124.5,
    lat = 33.5,
    legend_size = 2,
    # distance of one section of scale bar, in km
    distance_lon = 50,
    # height of the scale bar, in km
    distance_lat = 5,
    # distance between scale bar and units, in km
    distance_legend = 20,
    # units of scale bar
    dist_unit = "km",
    # add the north arrow
    orientation = TRUE,
    # length of N arrow, in km
    arrow_length = 60,
    # distance between scale bar & base of N arrow, in km
    arrow_distance = 50,
    # size of letter 'N' on N arrow, in km
    arrow_north_size = 10) +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())

# inset world map, from
# https://stackoverflow.com/a/62104396/1036500
library(maps)
library(magrittr)
library(maptools)
library(raster)

#Defining a general CRS
mycrs <- "+proj=longlat +datum=WGS84 +no_defs"

#Using the original maps package, then converting map into SpatialPolygons object
world <- maps::map("world", fill=TRUE,  plot = FALSE) %$%
  maptools::map2SpatialPolygons(., IDs=names,proj4string=CRS(mycrs))

#The resulting map has self intersection problems so any further operation reports errors; using buffers of width 0 is a fast fix
while(rgeos::gIsValid(world)==FALSE){
  world <- rgeos::gBuffer(world, byid = TRUE, width = 0, quadsegs = 5, capStyle = "ROUND")
}

#Dissolving polygon's limits
world <- raster::aggregate(world)

#Plotting. I add theme_void to your code to erase any axis, etc
worldplot <-
  ggplot() +
  geom_polygon(data = world,
               aes(x=long, y=lat,
                   group=group),
               fill='NA',
               color='black',
               size=0.1) +
  annotate("rect",
           xmin = 118,
           xmax = 137,
           ymin = 28,
           ymax = 47,
           fill = NA,
           colour = "red",
           size = 0.5
  ) +
  theme_void() +
  coord_fixed(1) +
  labs(x = "", y = "") +
  labs(x = NULL, y = NULL, title = NULL) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(panel.background = element_rect(fill = "white"),
        panel.border = element_blank(),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

# combine site map and inset map
library(cowplot)
ggdraw() +
  draw_plot(map_of_sites) +
  draw_plot(plot = worldplot,
            x = 0.45, # x location of inset placement
            y = -0.27, # y location of inset placement
            width = .6, # Inset width
            height = .75, # Inset height
            scale = 0.5 # Inset scale
  ) +
  theme(panel.background = element_rect(fill='white', colour="white"),
        plot.background = element_rect(fill='white', colour="white"))
```

\newpage

```{r}
#| label: tbl-sp-number
#| tbl-cap: Total number of stemmed points found in sites and the number of stemmed points we used in the research. We excluded broken and unclear artifacts.

here::here("analysis/data/raw_data/site-total-artifacts.csv") %>% 
  read_csv() %>% 
  knitr::kable(.)

```

\newpage

To analyze morphological change over time, we used previously developed chronologies that identify three phases in the Korean Late Paleolithic [@seong2015diversity; @Park_2013]. These chronologies were built on radiocarbon ages, and for sites without any, the presence/absence of blades, stemmed point blanks, and toolkit composition. Based on the existence of blade and micro blade and types of stemmed point blanks, we divided the Korean Late Paleolithic assemblages into three chronological phases: 1) stemmed points made out of flakes and no blades in assemblages, 2) stemmed points made out of blade or flake and the existence of blade in assemblages, and 3) stemmed points made out of blade and the existence of micro blade in assemblages (@tbl-korean-chronology). Applying this division is necessary because some sites such as Yonghodong, Goryeri, Jungmal, and Mungyeong have no radiocarbon ages due to various research limitations. Inferring ages for these sites by analogy to the technological sequences at sites with dates is important for making maximal use of the available archaeological data to increase our sample size. We arranged our `r total_number_of_assemblages` assemblages containing stemmed points in into these three phases to facilitate observations of change over time.

```{r}
#| label: tbl-korean-chronology
#| tbl-cap: Korean Late Paleolithic Chronology edited based on Seong [-@seong2015diversity] and Park [-@Park_2013]

here::here("analysis/data/raw_data/Korean-paleolithic-chronology.csv") %>% 
  read_csv() %>% 
  knitr::kable(.)

```

\newpage

Panel A of @fig-raw-materials-by-phase summarizes the amount of stemmed points in each chronological phase, and the distribution of raw materials across the phases. We exclude phase 1 from our analyses below because of the small sample size to compute correlation and CV. Shale is the dominant raw material in both phase 2 and 3. In phase 3 there is an increase in the proportions of porphyry, driven largely by finds from Yongsangdong, and other raw materials. Panel B in @fig-raw-materials-by-phase shows the count of stemmed points by raw material type at assemblages with more than five stemmed points. Much of the raw material diversity comes from isolated finds, with those assemblages with more than five points showing high homogeneity in raw materials.

```{r}
#| label: fig-raw-materials-by-phase
#| fig-cap: "A: Three chronological phases of the Korean Late Paleolithic, and frequecies of stemmed points and distribution of raw materials in each phase. B: Frequencies of stemmed points by assemblage and raw material, for assemblages with five or more stemmed points."
#| fig-height: 4

df_full_sitename_fct <- 
df_full_sitename %>% 
  mutate(SPstage.Raw_material = factor(SPstage.Raw_material))

#Create a custom color scale
library(RColorBrewer)
myColors <- 
  brewer.pal(length(unique(df_full_sitename_fct$SPstage.Raw_material)),
             "RdBu")
names(myColors) <- levels(df_full_sitename_fct$SPstage.Raw_material)
colScale <- scale_fill_manual(name = "Raw material",
                                values = myColors)

p1 <- 
ggplot(df_full_sitename_fct) +
  aes(SPstage.Stage,
      fill = SPstage.Raw_material) +
  geom_bar() +
  theme_bw() +
  labs(fill = "Raw material",
       x = "Chronological Phase") +
  theme(legend.position = "top",
        legend.key.size = unit(0.5, 'cm'), #change legend key size
        legend.key.height = unit(0.5, 'cm'), #change legend key height
        legend.key.width = unit(0.5, 'cm'), #change legend key width
        legend.title = element_text(size=8), #change legend title font size
        legend.text = element_text(size=6)
        ) +
colScale

p2 <- 
  df_full_sitename_fct %>%
  mutate(SPstage.Stage = paste0("Phase ", SPstage.Stage)) %>% 
  group_by(full_sitename) %>% 
  filter(n() > 5) %>% 
    ggplot() +
    aes(full_sitename,
        fill = SPstage.Raw_material) +
    geom_bar() +
    facet_wrap(~ SPstage.Stage) +
  theme_bw () +
  labs(fill = "Raw material",
       x = "Site") +
 theme(legend.position = "none",
       axis.text.x = element_text(size = 5)  ) +
  scale_x_discrete(guide = guide_axis(n.dodge= 2.3)) +
   colScale

library(cowplot)
plot_grid(p1, p2,
          ncol = 1,
          rel_heights = c(1.5, 0.75),
          labels = c("A", "B"))
```

\newpage

## Methods

Similar to the metric attributes from previous studies of cultural transmission and projectile points, i.e. maximum length and width [@bettinger1997evolutionary; @bettinger1999point; @eerkens2008cultural; @garvey2018current], we used morphological attributes on the stemmed points to examine variations in shape and the relationship between each attribute (@fig-labelled-schematic). We obtained our morphological attribute data from landmark analysis of digitized images of stemmed points. Compared to traditional caliper measurements, morphometric data yields more easily interpretable numerical and visual outcomes [@buchanan2010geometric; @thulman2012discriminating; @cardillo2016environment; @petvrik2018shape; @macleod2018quantitative; @okumura2019archaeology; @suarez2019life]. We did not include weight and thickness as these measurements were not available to us. For the landmark analysis, we put a total of 11 landmarks on the outline of each stemmed point and computed distances between pairs of landmark coordinates to derive attributes for our statistical analyses. The landmarks we recorded are described in @tbl-landmark-abbreviations and shown in @fig-labelled-schematic. We used the point tool in ImageJ [@schneider2012nih] to capture the landmarks from images of the artifacts, and export them as XY coordinate data for current and further analyses.

\newpage

```{r}
#| label: fig-labelled-schematic
#| fig-cap: Stemmed point from Yongsandong site with landmarks showing the attributes considered in this study and further research
#| out-width: "300px"

knitr::include_graphics(here::here("analysis/figures/hand-made-figures/ISKM-talk-2021-cut.png"))

```

\newpage

```{r}
#| label: tbl-landmark-abbreviations
#| tbl-cap: Description of artifact landmarks used in this study

here::here("analysis/data/raw_data/table-of-landmark-abbreviations.csv") %>% 
  read_csv() %>% 
  knitr::kable(.)

# to ensure we don't have any copy-paste errors 
# Correlation threshold
threshold_value_cor <- 0.5

# CV threshold
threshold_value <- 25
```

\newpage

Following @bettinger1999point we computed Pearson's correlation coefficients for all attributes on the stemmed points. The correlation of artifact attributes was proposed by @bettinger1999point to be a key indicator of indirect bias during cultural transmission. This is based on the assumption that artifact-making was transmitted in packages of traits taken from socially successful individuals who were role-models. In their work, @bettinger1999point interpreted correlation coefficient values of around `r threshold_value_cor` and higher as evidence of indirect bias, and lower values as evidence of guided variation. In this research, we examine correlation between attributes and explore other variables such as different phases, raw materials and assemblages.

We chose Coefficients of Variation (CV) to measure the variation among the measurements of the stemmed points. The CV is the ratio (%) of the standard deviation to the mean. It has been used as a robust statistical method to calculate the degree of standardization, precision, equality, homogeneity, etc. in various disciplines [@ng2006performance; @panichkitkosolkul2009improved; @panichkitkosolkul2013confidence]. In archaeological studies, CV has been applied to measure the variation between artifacts and to test hypotheses about cultural evolutionary processes, including distinguishing between different types of learning biases affecting cultural transmission [@eerkens2001techniques; @eerkens2005cultural; @schillinger2014copying; @garvey2018current]. Based on the Weber Fraction for visual measurement of line length, Eerkens and Bettinger [-@eerkens2001techniques] claim that over 57.7% of the CV is the result of random production and below 1.7% is the byproduct of using a scale or template.

Sample size can impact the accuracy of CV estimates [@toebe2018sample; @kelley2007sample; @vanpool2011quantitative]. In previous archaeological studies, sample sizes for CV values has varied from five to more than a thousand artifacts [@bettinger1999point; @garvey2018current; @costin1995standardization; @wierer2013variability; @kvamme1996alternative; @rivals2009new]. A variety of methods are available for improving the reliability of CV measurements when using small samples. For example, @vanpool2011quantitative proposed a "corrected CV", for a smaller sample (i.e. n\<25). Statistical research has resulted in several methods for computing confidence intervals on CVs to show uncertainty [@koopmans1964confidence; @miller1991asymptomatic; @sharma1994asymptotic; @mckay1932distribution; @vangel1996confidence; @curto2009coefficient; @banik2011estimating; @panichkitkosolkul2013confidence; @mahmoudvand2009two; @gulhar2012comparison]. Although we have a total of `r total_number_of_sp` artifacts, most of our analyses here involves comparisons of smaller subsets of this sample. For example to explore temporal and regional patterns we compared samples of well under 25 pieces. Given these relatively small sample sizes in our study, we used the corrected CV formula, which we have implemented here in a function in the R programming language for others to use. We also computed confidence intervals using method proposed by @sharma1994asymptotic to increase the credibility of the comparisons and be transparent about the precision and accuracy of our results [@kelley2007sample; @albatineh2014confidence].

@eerkens2001techniques presented some average CV values (%) of artifacts made from various materials (See Table 1 from @eerkens2001techniques). 
Among those various material artifacts, the most comparable artifacts to the stemmed points are Great Basin projectiles points, whose CV value ranged from 6-55, with an average of 22. Also, projectile points from the US Southwest have CV values ranged from 11 to 33 [@garvey2018current]. Unlike correlation, @bettinger1999point had vaguely mentioned about the threshold for CV values. Therefore, here we sat `r threshold_value` as a rough threshold to evaluate the transmission biases referring to the example of Great Basin projectile points. We calculated CV values of Rosegate points from Monitor Valley, which was speculated as the byproducts of indirect bias (raw data from table 5 of @bettinger1999point). With the exception of weight, the CV range for metric attributes on the points is 17.01 to 24.59. We chose higher side on the CV range of Rosegate points as the threshold of our research based on following reasons. 1) Variation is generated by small errors that are transmitted between individuals and the errors gets bigger through generations [@eerkens2005cultural]. The duration of transmission process for stemmed points is much longer (approx.30,000yrs) than the case of Rosegate (approx.500yrs in Fremont region) [@bischoff2021rosegate]. 2) The errors are also highly associated with raw materials. Some raw materials, such as clay, are easier to control while, less predictable and controllable materials such as stone are likely to have inflated CV values [@eerkens2001techniques]. We assume that Rosegate projectile points were made from more finer grained raw materials such as flint compared to raw materials for stemmed points such as shale, rhyolite, porphyry, etc. (@fig-raw-materials-by-phase). 3) Compared to normal CV, corrected CV has slightly higher values [@vanpool2011quantitative]. Based on the threshold, we premised that CV values that are lower than `r threshold_value` represent guided variation while CV values over `r threshold_value` reflect indirect bias.

All data preparation, analyses and visualization were computed in the R environment [@RCoreTeam]. Our R code and data, including the original artifact images, are fully and openly available in our compendium [@Marwick2018compendium] online at https://doi.org/10.17605/osf.io/eb8mx to enable transparency and reproducibility [@marwick2017computational].

# Results {.unnumbered}

## Correlation coefficients

Our results show positive correlation between attributes throughout the Late Paleolithic (Panel A of @fig-corr-analysis). Some relationships such as BL and ML have stronger correlation (i.e. darker-blue points) than others (i.e. lighter-blue or almost invisible points). To understand temporal patterns in the modes of cultural transmission, we grouped our assemblages into the three Korean Late Paleolithic chronological phases summarized in @tbl-korean-chronology. Among the three chronological phases, we excluded Phase 1 from our analysis because there are only two complete stemmed points from Yonghodong and Sachang that date to this phase.

The correlations between attributes became stronger over time as we can clearly see in the Panel D of @fig-corr-analysis. The median for Phase 2 is much lower than `r threshold_value_cor`, which @bettinger1999point interpreted as a threshold to evaluate transmission biases, while the Phase 3 median is close to the threshold. Based on this results, we assume that the earlier transmission for stemmed points was more likely the results of guided variation. On the other hand, Phase 3 could be more associated with indirect bias.

```{r}
#| label: fig-corr-analysis
#| fig-cap: "Correlation analysis by chronological phases in the Korean Late Paleolithic period. A. Correlation coefficient between attributes for all periods. B. Correlation coefficient between attributes for Phase 1. C. Correlation coefficient between attributes for Phase 2. The individual point represesnts correlation between two attributes. Some points are invisible due to their weak relationship. D. Correlation coefficient for the second and third chronological phases. The grey points represent correlation between two attributes."
#| fig-width: 7
#| fig-height: 5

library(corrr)

# correlation of all artifacts
df_sitename_corr <- 
  df_sitename %>%
  dplyr::select(-sitename, -SPstage.Stage, -SPstage.Raw_material) %>% 
  correlate() %>% 
  shave()

all_corr_plot <- rplot(df_sitename_corr, legend = FALSE)

df_sitename_corr_long <- 
  df_sitename_corr %>% 
    pivot_longer(-term)

# correlation of phase 2 artifacts
df_sitename_corr_p2 <- 
  df_sitename %>%
  dplyr::filter(SPstage.Stage == 2) %>%
  dplyr::select(-sitename, -SPstage.Stage, -SPstage.Raw_material) %>% 
  correlate() %>% 
  shave()

p2_corr_plot <- rplot(df_sitename_corr_p2, legend = FALSE)

df_sitename_corr_p2_long <- 
  df_sitename_corr_p2 %>% 
    pivot_longer(-term)

# correlation of phase 3 artifacts
df_sitename_corr_p3 <- 
  df_sitename %>%
  dplyr::filter(SPstage.Stage == 3) %>%
  dplyr::select(-sitename, -SPstage.Stage, -SPstage.Raw_material) %>% 
  correlate() %>% 
  shave()

p3_corr_plot <-  
  rplot(df_sitename_corr_p3) +
   theme(legend.position = c(0.9, 0.9),
         legend.key.size = unit(0.3, 'cm'), #change legend key size
        legend.key.height = unit(0.3, 'cm'), #change legend key height
        legend.key.width = unit(0.3, 'cm'), #change legend key width
        legend.title = element_text(size=6), #change legend title font size
        legend.text = element_text(size=8))

df_sitename_corr_p3_long <- 
  df_sitename_corr_p3 %>% 
    pivot_longer(-term)

# compare p2 and p3 correlations
compare_p2_and_p3_correlations_plot <- 
list(`phase 2` = df_sitename_corr_p2_long,
     `phase 3` = df_sitename_corr_p3_long) %>% 
  bind_rows(.id = "phase") %>% 
  ggplot() +
  aes(phase,
      value) +
  geom_boxplot() +
  ggbeeswarm::geom_quasirandom(size = 2,
                               alpha = 0.3) +
  theme_minimal() +
  labs(y = "Correlation coefficient") +
theme(axis.title.y = element_text(size = 13)) +
  # draw a line to show our threshold value
  geom_hline(yintercept = threshold_value_cor, colour = "red") +
  theme_bw()

plot_grid(plot_grid(
          plot_grid(
                    all_corr_plot,        
                    p2_corr_plot, 
                    p3_corr_plot,
                    ncol = 1,
                    labels = c("A", "B", "C")),
          compare_p2_and_p3_correlations_plot,
          ncol = 2,
          labels = c("", "D"),
          rel_heights = c(1, 2.5)))

```

One limitation of this aggregation of all stemmed points in our sample into each chronological phase is that the sample consists of a relatively large number of assemblages with only 1-2 stemmed points. These isolated finds are ambiguous with respect to a local tradition of artifact making, so to further investigate temporal change, we focus only the four assemblages that have more than five stemmed points (@fig-plot-cor-four-assemblage). We assume these four assemblages are more likely to represent a consistent, recurring way of making stemmed points than isolated finds, and thus more relevant for comparing modes of cultural transmission of artifact making.The four assemblages include three from the Suyanggae site (SYG) and one from the Yongsandong site (YS). There are more variation in correlation coefficient in Phase 2 while the median values for the two assemblages from Phase 3 are similar to each other and close to `r threshold_value_cor`. There is no clear pattern between assemblages. SYG1_2 is from Phase 2 but it has the higher median than the assemblage from the same site in the later phase.

```{r}
#| label: fig-plot-cor-four-assemblage
#| fig-cap: Correlation for assemblages with more than 5 stemmed points. Among the four assemblages, SYG1_2 and SYG6_2 are from Phase 2 of the Korean Late Paleolithic chronology and SYG1_3 and YS_3 are from Phase 3. The grey points represent correlation between two attributes.
#| fig-width:   4.5
#| fig-height:  3.5

#### plot cors by site with number of artifacts showing >5

cor_by_site_df_label_more_than_5 <-
  df_sitename %>%
  unite(col = "site_phase",
        c(sitename,
          SPstage.Stage),
        sep = "_") %>%
  dplyr::select(-SPstage.Raw_material) %>%
  dplyr::group_by(site_phase) %>%
  add_tally() %>%
  dplyr::filter( n > 5) %>%
  mutate(label = as.factor(paste0(site_phase,' (N = ', n, ")"))) %>%
  nest(-site_phase) %>%
  mutate(corrr = purrr::map(data, ~correlate(.x) %>%  
                                   shave() %>% 
                                  pivot_longer(-term))) %>% 
  unnest(corrr) %>% 
  filter()

#better arrangement for x axis for the plot
level_order <- c('SYG1_2', 'SYG6_2', 'SYG1_3', 'YS_3')

  ggplot(cor_by_site_df_label_more_than_5) +
  aes(x=factor(site_phase, level=level_order), y=value) +
  geom_boxplot() +
  ggbeeswarm::geom_quasirandom(size = 2,
                               alpha = 0.3) +
  theme_minimal() +
  labs(y = "Correlation coefficient",
       x = "Assemblage") +
theme(axis.title.y = element_text(size = 7.5)) +
  # draw a line to show our threshold value
  geom_hline(yintercept = threshold_value_cor, colour = "red") +
  theme_bw()

```

## Correlation between raw materials

To examine the impact of raw materials on the shape of stemmed points, we computed the correlation for different raw materials. We excluded raw materials that were used for less than three points. @fig-corr-raw-ass shows that the distribution of correlation coefficients is various depending on raw materials. Overall, the materials that used for more and equal to three points have high correlation coefficients. The stemmed points made from quartzite have highest correlation coefficients with narrow distribution. Acidic volcanic rocks are one of the highly correlated materials and individual correlations are skewed into two groups. Some of the correlations are even negatively related. Correlation values for porphyry stemmed points have a narrow distribution. Rhyolite stemmed points have wide distribution and some of the attributes are negatively related. Like acidic volcanic rocks, the correlation of shale stemmed points is also skewed into two groups. Tuff stemmed points have wide distribution with higher median value.

```{r}
#| label: fig-corr-raw-ass
#| fig-cap: Correlation for raw materials that were used for making more than one stemmed poitns. The grey points represent correlation between two attributes.

#| fig-width:7
#| fig-height:5

# correlation by site
# df_sitename_corr_site <- 
#   df_sitename %>%
#   group_by(sitename) %>% 
#   filter(n() > 5) %>% 
#   dplyr::select(-sitename, 
#                 -SPstage.Stage) %>% 
#   nest(-sitename) %>% 
#   mutate(corrr = purrr::map(data, ~correlate(.x) %>%  
#                                    shave() %>% 
#                                   pivot_longer(-term))) %>% 
#   unnest(corrr) %>% 
#   filter()
# 
# correlations_by_site <- 
#   ggplot(df_sitename_corr_site) +
#   aes(sitename,
#       value) +
#   geom_boxplot() +
#   ggbeeswarm::geom_quasirandom(size = 2,
#                                alpha = 0.3) +
#   theme_minimal() +
#   labs(y = "Correlation coefficient",
#        x = "Site") +
# theme(axis.title.y = element_text(size = 4)) +
#   # draw a line to show our threshold value
#   geom_hline(yintercept = threshold_value_cor, colour = "red") +
#   theme_bw()

# correlation by raw material
df_sitename_corr_raw_material <- 
  df_sitename %>%
  group_by(SPstage.Raw_material) %>% 
  filter(n() >= 3) %>% 
  dplyr::select(-sitename, 
                -SPstage.Stage) %>% 
  nest(-SPstage.Raw_material) %>% 
  mutate(corrr = purrr::map(data, ~correlate(.x) %>%  
                                   shave() %>% 
                                  pivot_longer(-term))) %>% 
  unnest(corrr) %>% 
  filter()

# correlations_by_raw_material <- 
  ggplot(df_sitename_corr_raw_material) +
  aes(SPstage.Raw_material,
      value) +
  geom_boxplot() +
  ggbeeswarm::geom_quasirandom(size = 2,
                               alpha = 0.3) +
  theme_minimal() +
  labs(y = "Correlation coefficient",
       x = "Raw material") +
theme(axis.title.y = element_text(size = 4)) +
  # draw a line to show our threshold value
  geom_hline(yintercept = threshold_value_cor, colour = "red") +
  scale_x_discrete(guide = guide_axis(n.dodge = 2))+
  theme_bw()

```

## Coefficient of variation

```{r}
#| label: prepare-for-cv-calcs

# Coefficient of variation (CV): the ratio of the standard deviation divided by mean

# cv <- function(x, ... ) sd(x, ...) / mean(x, ...) * 100

# use this one that is optimized for small samples, returns value 0-1, not %. 
# We multiply 100 (%) later, altogether with confident interval.

mccv <- function(x, ... ) sd(x, ...) / mean(x, ...) * (1 + 1/(4*length(x)))

# BM: this is the cvCI function from https://github.com/stamats/MKmisc, I 
# removed all the non-sharma methods and replaced CV() with mccv() above

sharma_cvCI <- function(x, 
                 conf.level = 0.95, 
                 method = "sharma", 
                 na.rm = FALSE){
    stopifnot(is.numeric(x))

    if(length(conf.level) != 1)
        stop("'conf.level' has to be of length 1 (confidence level)")
    if(conf.level < 0.5 | conf.level > 1)
        stop("'conf.level' has to be in [0.5, 1]")

    alpha <- 1 - conf.level
    zquant <- qnorm(1-alpha/2)
    cv <- mccv(x = x, na.rm = na.rm)
    Infos <- NULL
    if(na.rm) x <- x[!is.na(x)]
    n <- length(x)

     # Sharma and Krishna 1994
      CI.lower <- 1/(1/cv + zquant/sqrt(n))
      CI.upper <- 1/(1/cv - zquant/sqrt(n))
      METH <- "Sharma and Krishna (1994) confidence interval"

    CI <- matrix(c(CI.lower, CI.upper), nrow = 1)
    rownames(CI) <- "CV"
    colnames(CI) <- c(paste(alpha/2*100, "%"), paste((1-alpha/2)*100, "%"))
    attr(CI, "conf.level") <- conf.level
    names(cv) <- "CV"

    return(structure(list("estimate" = cv, "conf.int" = CI, "Infos" = Infos,
                          "method" = METH),
                     class = "confint"))
}
```

```{r}

# we've got DF a data frame with cols as variables
# rows and specimens, and values as dimensions
# need to add a column of site ID

DFCV <- map_dbl(DF, mccv)

# GP's test for comparing CV vs corrected CV
# mock_test <-
#   readr::read_csv(here::here("analysis/data/raw_data/mock_test.csv"))
# mock_test_cv <-map_dbl(mock_test, cv)
# mock_test_mccv <- map_dbl(mock_test, mccv)

#Same as map but better view for summary
cv_all <- map_df(DF, mccv)
#arrange the data properly
cv_data_all <- gather(cv_all)

# to install MKmisc
## Install package BiocManager
# install.packages("BiocManager")
## Use BiocManager to install limma
# BiocManager::install("limma")
# install.packages("MKmisc")

library(MKmisc)
library(purrr)
# CV value
sharma_cv <- function(x,...) sharma_cvCI(x, conf.level = 0.95, method = "sharma", na.rm = FALSE)$estimate * 100

#CV sharma interval values
sharma_int_low <- function(x,...) sharma_cvCI(x, conf.level = 0.95, method = "sharma", na.rm = FALSE)$conf.int[1] * 100

sharma_int_high <- function(x,...) sharma_cvCI(x, conf.level = 0.95, method = "sharma", na.rm = FALSE)$conf.int[2] * 100

all_cv_sharma <- map_dbl(DF, sharma_cv)
all_low_sharma <- map_dbl(DF, sharma_int_low)
all_high_sharma <- map_dbl(DF, sharma_int_high)

#create data frame
cv_data_sharma <- data.frame(all_cv_sharma, all_low_sharma, all_high_sharma)

#add column name
library(tibble)
cv_data_sharma_named <- rownames_to_column(cv_data_sharma, var = "attribute")

# for in-text numbers
max_cv <- round(max(cv_data_sharma_named$all_cv_sharma), 1)
min_cv <- round(min(cv_data_sharma_named$all_cv_sharma), 1)

```

```{r}
#| label: fig-plot-attributes
# Box plot for each attribute
# mutate data to make plot with ggplot
data_box <- gather(DF)
# make box plot for each attributes
plot_raw_measurements_for_attributes <- 
ggplot(data_box,
       aes(key, value)) +
  geom_boxplot() +
  geom_beeswarm(alpha = 0.3,
                size = 0.8,
                cex = 0.7) +
  ylab("Length (cm)") +
  xlab("") +
  theme_bw()
```

Panel A of @fig-plot-cv-all-attributes shows a relatively wide distribution of body length (BL) and maximum length (ML) while the other attributes are more similar and shorter (under 2.5 cm). Among these attributes, stem width (SW) is the smallest, regardless of overall size of the stemmed points.

Panel B of @fig-plot-cv-all-attributes shows that CV values for all attributes are distributed from `r min_cv` to `r max_cv`. Compared to the CVs of Great Basin projectile points [@eerkens2001techniques], the average CV for the Korean artifacts is higher. Unlike the distribution of the linear measurements, the CVs of body length and maximum length have the lowest values. The lower CV of maximum length indicates they are highly standardized with narrowest confidence intervals. Tang and stem related attributes (e.g. SL, SW, TL, TW) are less standardized with higher CV values and wider confidence intervals than other attributes. Overall, our results suggests that transmission of stemmed points could be more influenced by guided variation.

```{r}
#| label: fig-plot-cv-all-attributes
#| fig-cap: A. Boxplot for each attribute measured on the stemmed points. The black bold horizontal lines in the middle of boxes represent the median value. The dots represent individual artifacts. B. The points and blue colored numbers represent CV values for each attribute measured on the stemmed points. 95% confidence intervals for the CVs were computed using Sharma and Krishna's method. The vertical lines indicate the confidence intervals.

#| fig-width:7
#| fig-height:4

plot_cv_all_attributes <- 
ggplot(cv_data_sharma_named,
       aes(attribute,
           all_cv_sharma)) +        # ggplot2 plot with confidence intervals
  geom_point() +
  geom_errorbar(aes(ymin = all_low_sharma, 
                    ymax = all_high_sharma), 
                width = 0.2) +
  ylab("CV") +
  xlab("Artifact measurement") +
  geom_text(aes(label = round(all_cv_sharma,1)), 
            col="blue", 
            hjust = -0.3, 
            size = 3) +
  # draw a line to show our threshold value
  geom_hline(yintercept = threshold_value, colour = "red") +
  theme_bw()

library(cowplot)
plot_grid(plot_raw_measurements_for_attributes,
          plot_cv_all_attributes,
          ncol = 1, 
          labels = "AUTO")
```

Similar to Panel D of @fig-corr-analysis, we divided the stemmed points into three chronological phases and computed CV to examine temporal patterns in the modes of cultural transmission. Again, we excluded stemmed points for Phase 1 due to the limited number of artifacts.

@fig-plot-cv-over-time shows that the directional trend in CV values of artifact attributes is complicated. Half of the attributes are below our threshold value of `r threshold_value`, and half are above. Only body length crosses the threshold value, changing from \<`r threshold_value` to \>`r threshold_value` over time, indicating a shift from indirect bias to guided variation. For those attributes where the CV is \<`r threshold_value`, the trend is decreasing CV values from Phase 2 to Phase 3 for maximum width and tang length, indicating increasing influence of indirect bias for these attributes. For stem length, stem width and tang length, CV values remain above `r threshold_value`, and show a slight trend to increase over time, suggesting increased influence of guided variation for these attributes.

```{r}
#| label: fig-plot-cv-over-time
#| fig-cap: CV values for the second and third chronological phases in the Korean Late Paleolithic period. The points represent CV values for each attribute. The vertical lines indicate the confidence intervals.
#| fig-width:   4.5
#| fig-height:  4

#-- temporal patterns ----------------------------------------------------------------
# Make a table of CVs for all variable grouped by stage, according to
# stemmed point chronology
# Exclude the first phase that has only two artifacts

cv_by_stage_df  <-
  df_sitename %>%
  dplyr::select(-sitename,
         -SPstage.Raw_material) %>%
  group_by(SPstage.Stage) %>%
  add_tally() %>%
  dplyr::filter(n >= 5) %>%
  dplyr::select(-n) %>%
  pivot_longer(-SPstage.Stage,
               names_to = "variable",
               values_to = "value") %>%
  group_by(SPstage.Stage, variable) %>%
  dplyr::summarise(cv_by_site =   sharma_cv(value),
            cv_low_sharma =  sharma_int_low(value),
            cv_high_sharma = sharma_int_high(value))

# facet plot for two phases
cv_by_stage_df %>%
  ggplot(aes(x = factor(SPstage.Stage),
             y = cv_by_site)) +
  geom_point() +
  geom_errorbar(aes(ymin = cv_low_sharma,
                    ymax = cv_high_sharma),
                width = 0.2) +
  xlab("Phase") +
  ylab("Coefficient of Variation for all Attributes") +
    # draw a line to show our threshold value
  geom_hline(yintercept = threshold_value, colour = "red") +
  facet_wrap( ~ variable, scales = "free_y") +
  theme_bw()
```

Following the previous studies of [@bettinger1997evolutionary; @bettinger1999point; @eerkens2008cultural; @mesoudi2008cultural; @garvey2018current], we only compared CV values of attributes, body length, maximum length, mid width, and tang width for the assemblages that have more than five stemmed points. @fig-plot-cv-four-assemblage shows that CV values for the four assemblages are mostly slightly below `r threshold_value`. Among the assemblages, SYG1_2 has higher CV values as well as the widest ranges of confidence intervals. It is interesting because SYG1_2 has the strong correlation between attributes (@fig-plot-cor-four-assemblage). Perhaps it could be related to its small sample size, relative to the other assemblages compared here. Overall we see only subtle changes in CV values from Phase 2 to Phase 3 in @fig-plot-cv-four-assemblage. This suggests that the complex directional trends in @fig-plot-cv-over-time might be best interpreted as noise in an overall signal of guided variation, rather than substantial changes in the type of bias dominating cultural transmission.

```{r}
#| label: fig-plot-cv-four-assemblage
#| fig-cap: CV values of four attributes from assemblages with more than 5 stemmed points. Among the four assemblages, SYG1_2 and SYG6_2 are from Phase 2 of the Korean Late Paleolithic chronology and SYG1_3 and YS_3 are from Phase 3. The points represent CV values for each attribute. The vertical lines indicate the confidence interval.
#| fig-width:   4.5
#| fig-height:  4

#### plot CVs by site with number of artifacts showing >5

cv_by_site_df_label_more_than_5 <-
  df_sitename %>%
  unite(col = "site_phase",
        c(sitename,
          SPstage.Stage),
        sep = "_") %>%
  dplyr::select(-SPstage.Raw_material) %>%
  dplyr::group_by(site_phase) %>%
  add_tally() %>%
  dplyr::filter( n > 5) %>%
  mutate(label = as.factor(paste0(site_phase,' (N = ', n, ")"))) %>%
  ungroup() %>%
  dplyr::select(-n, -site_phase) %>%
  pivot_longer(-label,
               names_to = "variable",
               values_to = "value") %>%
  dplyr::group_by(label, variable) %>%
  dplyr::summarise(cv_by_site =   sharma_cv(value),
            cv_low_sharma =  sharma_int_low(value),
            cv_high_sharma = sharma_int_high(value)) %>%
  dplyr::filter(variable %in% c("ML", "BL","MW", "TW"))

cv_by_site_df_label_more_than_5 %>%
  ggplot(aes(x = factor(variable),
             y = cv_by_site)) +
  geom_point() +
  geom_errorbar(aes(ymin = cv_low_sharma,
                    ymax = cv_high_sharma), width = 0.2) +
    # draw a line to show our threshold value
  geom_hline(yintercept = threshold_value, colour = "red") +
  xlab("Attribute") +
  ylab("Coefficient of Variation on Attributes") +
  facet_wrap( ~ label, scales = "free_y") +
  coord_cartesian(ylim = c(15,60)) +
  theme_bw()

```

## Variation Between Raw Materials

We examined the relationship between CV values and raw materials to test the hypothesis that the shape of stemmed points was highly dependent on raw material. We excluded raw materials that were used for less than three points. @fig-plot-cv-by-raw-material shows that quartzite, the raw material for the stemmed points at Sachang, has the highest CV values, and also has wide conference intervals. Other raw materials are generally stable in variation throughout the attributes.

```{r}
#| label: fig-plot-cv-by-raw-material
#| fig-cap: CV values for raw materials that were used for making more than one stemmed point. The points represent CV values for each attribute. The vertical lines indicate the confidence intervals.
#| fig-width:   8.5
#| fig-height:  4

# CV by raw materials
### Make a table of CVs for all variable grouped by raw materials
cv_by_raw_materials <-
  df_full_sitename %>%
  dplyr::select(-SPstage.Stage, -lat_dd, -long_dd, -sitename, -full_sitename) %>%
  group_by(SPstage.Raw_material) %>%
  add_tally() %>%
  dplyr::filter( n >= 3) %>%
  dplyr::select(-n) %>%
  pivot_longer(-SPstage.Raw_material,
               names_to = "variable",
               values_to = "value") %>%
  group_by(SPstage.Raw_material, variable) %>%
  dplyr::summarise(cv_by_materials =   sharma_cv(value),
            cv_low_sharma =  sharma_int_low(value),
            cv_high_sharma = sharma_int_high(value))

######CV for per raw materials (n=X) with the full site name : label
cv_by_raw_materials_df_label <-
  df_sitename %>%
  group_by(SPstage.Raw_material) %>%
  add_tally() %>%
  dplyr::filter( n > 1) %>%
  mutate(label = as.factor(paste0(SPstage.Raw_material,' (N = ', n, ")"))) %>%
  dplyr::select(-n) %>%
  nest(-SPstage.Raw_material, -label) %>%
  mutate(cv_by_site = purrr::map(data, ~map_df(.x, mccv))) %>%
  dplyr::select (-cv_by_site)

######add label to main dataframe
cv_plot_raw_materials_label <- cv_by_raw_materials %>%
  left_join (cv_by_raw_materials_df_label) %>%
  dplyr::select (- data,-SPstage.Raw_material)

###### facet plot
cv_plot_raw_materials_label %>%
  ggplot(aes(x = variable, y = cv_by_materials)) +
  geom_point() +
  geom_errorbar(aes(ymin = cv_low_sharma,
                    ymax = cv_high_sharma), width = .2) +
    # draw a line to show our threshold value
  geom_hline(yintercept = 25, colour = "red") +
  xlab("Raw Materials") +
  ylab("Coefficient of Variation on Attributes") +
  facet_wrap( ~ label, scales = "free_y") +
  coord_cartesian(ylim = c(0,250)) +
  theme_bw() 
```
# Discussion

To investigate the social contexts of technological transitions represented by the emergence of stemmed points during the Korean Late Paleolithic, we applied concepts of cultural transmission. We asked three questions to examine the dominant mode of cultural transmission and the transmission process over time: what is the dominant mode of cultural transmission for technological innovation in the Korean Late Paleolithic? Do the modes of cultural transmission vary over time? and Do the modes of cultural transmission vary over space? We focused on two transmission biases: guided variation (socially isolated groups made stemmed points through trial and error) and indirect bias (socially connected groups whose knowledge of stemmed points derived from copying others). Following @bettinger1999point, we used `r threshold_value_cor` as a threshold of correlation coefficients to evaluate the two transmission biases. Correlation coefficients below `r threshold_value_cor` are indicative of guided variation and higher values indicate indirect bias. Since there is no clear criteria for CV, we set roughly `r threshold_value` as our threshold of the CV value to decide between these two types of transmission biases, based on prior work on Great Basin Projectiles [@eerkens2001techniques]. If the CV value is lower than `r threshold_value`, we interpreted this as indicating indirect bias while values higher than `r threshold_value` were interpreted as guided variation.

Our correlation analyses show that the correlation coefficients of stemmed points are all positive but mostly under `r threshold_value_cor`, showing the possibility of guided variation as their main transmission bias. Some correlations such as body length and maximum length have strong relationship while other correlations such as tang length and body length are much weaker. The correlation coefficients increase from chronological Phases 2 to 3 (Panel D of @fig-corr-analysis), but there is no clear pattern and the values vary depending on assemblages (@fig-plot-cor-four-assemblage). The CV values for all stemmed point attributes are mostly over `r threshold_value`, except for maximum length, which is just below. Change in CV values for over time in the Korean Late Paleolithic period is complex, with no clear directional changes (@fig-plot-cv-over-time). Comparing chronological phases 2 and 3, we found that there are only minor differences between the two phases. Applying the Modified Signed-Likelihood Ratio Test (MSLR) to test for the equality of CVs [@smallwood2022using; @krishnamoorthy2014improved], we found no statistically significant changes in the CV values of any attributes (@tbl-cv-over-time) between the two phases. In exploring individual assemblages that contain multiple stemmed points we found confirmation of minimal change in CV values over time (@fig-plot-cv-four-assemblage).

```{r}
#| label: tbl-cv-over-time
#| tbl-cap: "Summary of significance tests for CV values between the two chronological phases"

####### significant test between the two phases ######
library(cvequality)

# created nested tibble, nest by variable
df_sitename_nested <-
  df_sitename %>%
  dplyr::select(-sitename, -SPstage.Raw_material) %>%
  group_by(SPstage.Stage) %>%
  add_tally() %>%
  dplyr::filter(SPstage.Stage > 1) %>%
  dplyr::select(-n) %>%
  pivot_longer(-SPstage.Stage,
               names_to = "variable",
               values_to = "value") %>%
  ungroup() %>%
  group_by(variable) %>%
  nest()

# apply mslr test to each variable (row with a tibble)
# Krishnamoorthy and Lee’s (2014) modified signed-likelihood ratio test
# this is a table we can show in the text
table_mslr_test_results_for_phase_differences <-
  df_sitename_nested %>%
  mutate(mslr_result = map_df(data, ~mslr_test(nr = 1e4,
                                               .$value,
                                               .$SPstage.Stage))) %>%
  unnest(mslr_result) %>%
  dplyr::select(-data) %>% 
  # round numeric columns to 3 decimal places
  mutate(across(where(is.numeric), round, 3)) %>% 
  arrange(desc(p_value))

names(table_mslr_test_results_for_phase_differences) <- 
  c("Variable", "MSLR statistic", "p value")

knitr::kable(table_mslr_test_results_for_phase_differences,
             digits = 3)

```
In exploring geographical variation we found no clear pattern variation in both correlation coefficients (@fig-plot-cor-four-assemblage) and CV values (@fig-plot-cv-four-assemblage) between assemblages with more than five stemmed points. We found that each assemblage has different distribution patterns of correlation coefficients and CV values. SYG1_2 and SYG1_3 are from the same site but different chronological periods and correlations decreases over time. It is likely that correlation or overall shape can be more depending on individual knappers. We also found that some attributes such as maximum width and maximum length have stronger correlation while the others including stemmed length and body length have lower or even negative correlation. We assume that certain attributes were more carefully transmitted than others. We observed that CV values of body length, maximum length, mid width, and tang width are generally lower than tang length, stem length, and stem width, similar to what is evident in @fig-plot-cv-all-attributes. The attributes with lower CV values could be transmitted in a social context dominated by indirect bias, while the higher CV attributes could be influenced more by specific manufacturing and maintenance situations. We assume that the lower CV values of the four attributes could imply that those four attributes are closely associated with the projectile function of the tool so that the knappers wanted to standardize the shape. For example, body length is related to penetration, durability, hardness, and rejuvenation of a projectile point and tang width is connected to wound damage, penetration, durability and hardness [@odell1986experiments; @shea2001experimental; @cheshier2006projectile; @yaroshevich2016projectile; @bebber2017developing; @wood2018wound]. This can be contrasted with tang length, stem length, and stem width, which are away from the impact end of the artifact. These were allowed to vary more freely between points, perhaps to be able to accommodate shafts of different types of wood with varying properties of strength, flexibility, and weight.

The most striking differences in both correlation coefficients and CV values is 
in comparisons of points grouped by raw materials (@fig-corr-raw-ass, @fig-plot-cv-by-raw-material). All distributions are different and there is no clear pattern, indicating that raw materials strongly influence stemmed points. We expected that stemmed points with higher correlation coefficients would have lower CV values, but the actual results do not match our expectations. For example, artifacts made from acidic quartzite have highest correlation but their CV values are also much higher with wider confidence intervals, than all other raw materials. This conflicting results might be related to the relatively unpredictable knapping quality of quartzite. These grains can redirect knapping forces in unexpected directions, resulting in unintended flake removals. This makes is harder for the knapper to repeatedly produce the same size and shape end product. This hints at the possibility that physical properties of the stone artifact raw materials were a key mechanism in determining shape and size variability. However with only three artifacts made from quartzite in our sample, further analysis is needed to make robust conclusions about the influence of raw materials here.

The results of this study have implications for determining whether these novel technologies originated outside of the Korean Peninsula or if they were locally developed independently.??

# Conclusion

In this research, we investigated the social context of technological innovation in the Korean Late Paleolithic. Following previous studies of applying cultural transmission to the introduction of bow and arrow technology in the Great Basin [@bettinger1997evolutionary; @bettinger1999point; @eerkens2008cultural; @garvey2018current], we investigated two transmission biases, guided variation (trial and error) and indirect bias (copying a model). We proposed two scenarios for explaining the introduction of new technology: socially isolated groups that made stemmed points through trial and error (guided variation) or socially connected groups whose knowledge of stemmed points derived from copying others that they regularly came in contact with (indirect bias). We asked three questions: what is the dominant mode of cultural transmission for technological innovation in the Korean Late Paleolithic? Do the modes of cultural transmission vary over time? And do the modes of cultural transmission vary over space?

We conclude that the dominant mode of cultural transmission for the earliest stemmed points was guided variation (trial and error). This is indicated by mostly low correlation coefficients of less than `r threshold_value_cor` and relatively high CV values of around and above `r threshold_value`. A social context that favored trial and error for artifact making is consistent with Seong's *in situ* model for the appearance of stemmed points in Korea. This model proposes that people developed stemmed points by experimenting with existing blade technologies. Our results support experimentation and trial and error as important processes in the cultural transmission of this artifact technology.

The correlation coefficients suggest a change in the dominant bias of cultural transmission over time, with values around `r threshold_value_cor` in chronological phase 3, after 25 ka. This is in the same range of correlation coefficients that Bettinger and Eerkens interpreted as indirect bias. We also confirm a slight decrease in CV values in four attributes. This results may suggest a shift in the social context of transmission of tool-making skills away from trial and error and towards increased reliance on copying a model. We might speculate based on the evolutionary ecological assumption that cooler conditions of the Last Glacial Maximum promoted greater integration of social networks to buffer risks of resource failure [@fitzhugh2001risk]. With more frequent contacts and connections with members of the social network, there may have been more opportunities to learn point-making directly from socially successful individuals.

In comparing assemblages with more than five stemmed points, we found complex variation in CV values that may suggest mixtures of biases in the transmission of tool-making techniques. We did not find strong support for distinctive forms of cultural transmission of tool-making at different assemblages. Correlation coefficient values across the assemblages affirm the overall chronological trend indicating indirect bias in phase three. Raw materials appear to be important in driving this trend, with acid volcanic rocks and porphyry becoming more abundant in phase 3, and dominating assemblages at Yonsangdong that date to this period. One possibility is that these raw materials were part of an adaptive shift during the Last Glacial Maximum, perhaps the cooler climate favored an expanded range for searching for raw materials, leading to a higher diversity. Similarly, change in occupation and mobility patterns may have resulted in more frequent contacts and connections with members of social networks [@park2022did]. This may have provided more opportunities to learn point-making directly from socially successful individuals, and increased the contribution of indirect bias during cultural transmission during phase 3.

# Acknowledgements

<!-- The following line inserts a page break  -->

# CRediT authorship contribution statement

Gayoung Park: Conceptualization, Software, Validation, Formal analysis, Methodology, Resources, Data curation, Writing - original draft, Writing - Review & Editing, Visualization, Project administration.

Ben Marwick: Software, Validation, Formal analysis, Investigation, Writing - Review & Editing, Visualization, Supervision

\newpage

# References {.unnumbered}

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

::: {#refs}
:::

\newpage

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```